### 2024-04-24 05-conclusiones.qmd
render paper

---
author: "Equipo EDUMER"
format: pdf
editor: visual
---

## Conclusion

-   Efecto ilustrador de la educación? (esfuerzo -\> redistribución)