### 2024-04-25 02-antecedentes.qmd
update references

---
author: "Equipo EDUMER"
format: pdf
editor: visual
---

## Justification of inequality

Research on social stratification beliefs, which explore individual perceptions of who deserves what and why [@kluegel_beliefs_1987], highlights that people’s explanations and justifications of social inequality are closely tied to their judgments of deservingness. The influence of ideologies [@wegener_dominant_1995] and cultural schemas [@homan_being_2017] is pivotal in shaping these explanations by offering symbolic representations that frame societal structures and expectations. While significant attention has been paid to wage inequality, income distribution, and payment differentials in the literature [@castillo_legitimacy_2011; Evans et al., 2010; @jasso_how_1999 ; @shariff_income_2016],  there has been less examination of public beliefs about which life domains should be governed by market relations and even less about children’s acceptance or rejection of these market principles. This oversight is notable given the extensive encroachment of market logic into public goods, welfare policy, and social services over the past five decades [@centeno_arc_2012; @harvey_breve_2015], affecting areas such as pensions, health services, and education. The justification of social inequality based on marketype criteria has been conceptualized as the individuals’ adherence to one specific justice evaluation, the market justice, that is, affording legitimacy to the allocation of goods and services based on prices and individuals’ ability to pay [@boltanski_new_2005; @lane_market_1986; @streeck_citizens_2012].

Robert E. Lane proposed the underpinnings of market justice, which he differentiated from political justice. For him, “it is the genius of the market to stimulate wants without at the same time stimulating a sense of deserving more than one gets” [-@lane_market_1986; p. 384]. Following the theory of relative deprivation –a social phenomenon arising when individuals cannot afford what most others in their environment can (Merton 1950)– Lane notes that, in market settings, social comparisons are more likely to motivate increased effort rather than feelings of acute injustice because individuals attribute outcomes to their actions. Although empirical research has shown that, contrary to Lane’s observation, relative deprivation, even in market settings, produces feelings of dissatisfaction, anger, and resentment that might motivate forms of collective action such as protests and revolt [@greitemeyer_subjective_2016; @mishra_subjective_2015; @smith_relative_2012; @power_deprivationprotest_2018; @power_relative_2020], the conceptualization of market justice has been since then closely coupled to the merit principle for allocating outcomes: the claim that the unequal levels of well-being individuals enjoy ought to be, to some extent, a function of their talents and efforts, regardless of their needs or membership, the two latter being the realm of political justice and its closely coupled principles of need –allocating outcomes to those who require them most– and equality –allocating the same outcome to everyone–, which have been at the center of welfarism (see Wilson 2003). The underlying notion of market justice also resonates in its application to welfare regimes, providing a framework for understanding the varied global approaches to managing social services.

The management of social services manifests in varying approaches across nations, with substantial differences in funding and delivery methods [@jensen_worlds_2008; @stoy_worlds_2014]. Nordic countries, for example, predominantly employ public agencies to produce and provide social services, funding these through collective taxation and offering them in kind to the majority of citizens. This system prioritizes political justice, placing it above market mechanisms in accessing services. In contrast, other countries rely more heavily on for-profit entities and private funding, where service distribution depends mainly on individual financial capacity of paying user fees, highlighting the influence of market justice in service allocation. The trend toward marketization of welfare services has been growing since the 1980s [@salamon_marketization_1993], and this shift is increasingly evident even in countries where market solutions have traditionally had a minor role in social policy [@sivesind_changing_2017].

The question arises whether adults and children justify unequal access to welfare services based on market justice principles. Influenced by theories of policy feedback, which suggest that social welfare policies can reinforce (positive feedback) or undermine (negative feedback) previous policy trajectories [@fernandez_positive_2013; @pierson_increasing_2000; @weaver_paths_2010], citizens’ beliefs about market justice are likely also shaped by the institutional and social contexts they encounter. Indeed, the justification of inequality in access to essential services like education and health, based on one’s ability to pay, shows significant variation across countries, as demonstrated by international surveys, although they do not usually include children in their samples. For instance, the 2019 International Social Survey Program (ISSP) provides insights into how adults perceive inequality in accessing welfare services. Figure 1 illustrates the variation in agreement levels regarding whether it is just for individuals with higher incomes to purchase better education or healthcare. Notably, in 18 of the 29 countries analyzed, there is a greater justification for market inequality in healthcare access than in education. Nevertheless, the general sentiment typically ranges from 'somewhat unjust' to 'neither just nor unjust, with mixed feelings.

```{r}
#| fig-cap: "Average of market justice preferences by country"
```

While contemporary policy developments have increasingly embraced the marketization of social welfare in areas such as education, pensions, and health services, questions about public acceptance of these changes persist. @lindh_public_2015 analysis of ISSP 2009 data from 17 OECD countries reveals a general lack of support for market-based distribution of social services, suggesting widespread disapproval of market stratification of essential services. This finding is corroborated by @soler-martinez_concerns_2023 research from Latinobarómetro 2020 across 18 Latin American countries, where concerns about health and education access predominated over income inequality. These results indicate that reforms toward welfare marketization are typically driven by elite political decisions rather than grassroots demand.

Despite high-income inequality and limited social mobility in Latin America, there is a prevalent belief that individuals are solely responsible for their economic outcomes, a view that varies across the region [@bucca_merit_2016; @chong_mystery_2008; @torche_intergenerational_2014; @salgado_inequality_2023]. The reliance on private welfare providers and widespread user fees [@molyneux_neoliberal_2008] adds complexity to this context. Yet, research on children’s justification of market-based inequalities in accessing welfare services remains limited, especially in Latin America, highlighting a significant gap in understanding how younger generations view market-based access to welfare and whether these views are associated with their meritocratic beliefs.

## Meritocracy

The concept of meritocracy frequently appears nowadays when analyzing cultural determinants of social inequalities. In general, it is mentioned as a value associated with justice, as it would link efforts and talents with rewards in an equitable manner. This normative sense is quite far from its original formulation by @young_rise_1958 in the satirical novel "The Rise of Meritocracy", where it ironically represented a mechanism for reproducing the inequalities of origin. The meritocratic ideal had remained relatively unchallenged until a series of recent publications turned into its potential consequences for maintaining social inequality. Perhaps one of the most recent sources in this line is Michael Sandel's "The Tyranny of Merit", where he strongly questions the implications of carrying out the principle of merit in societies that do not guarantee equal opportunities and that generate a feeling of scarce recognition and appreciation of those who receive lesser rewards: "In society's eyes, and perhaps also their own, their work no longer signified a valued contribution to the common good." [@sandel_tyranny_2020, pp.].

Empirical research on meritocracy has increased along with the philosophical-normative discussion on meritocracy in recent years. Particularly from a sociological perspective, meritocracy has been used in research on social mobility to characterize societies with low mobility that threaten the meritocratic ideal [@goldthorpe_myth_2003]. More recently, sociology and social psychology research has attended to the subjective aspects vis-a-vis beliefs in meritocracy. The label of beliefs in this realm covers a series of areas, such as attitudes, perceptions, and preferences (Castillo et al), whereby most of the link this subjective dimensions to individual socio-structural factors and  context-level determinants. For instance, some studies have analyzed how those with greater privileges believe more in meritocracy (Reynolds & Chan 2014), how greater economic inequality increases meritocratic beliefs [@mijs_paradox_2021], and how larger inequality affects meritocratic beliefs [@morris_representing_2022]. Based on these findings, a research agenda has been reinforced on the legitimizing role of meritocracy, in line with previous studies using the concept of a just world (Lerner, Dalbert) and the theory of system justification (Jost & Major).

How do meritocratic beliefs legitimize inequalities? Empirical studies have used experiments and surveys to address this question. For instance, the evidence suggests that just world beliefs correlate negatively with support for redistributive compensation systems [@frank_performance_2015]. Conversely, individuals tend to support redistribution when they believe that the disadvantaged lack the opportunities to succeed [@evans_strong_2018]. @almas_cutthroat_2020 found that in a relatively unequal society (the United States), the highly educated accept inequality significantly more than the less educated because they perceive inequality as justifiable owing to differences in productivity (i.e., merit), whereas in a relatively equal society (Norway), the less educated accept inequality more, but not significantly more than the highly educated because meritocratic values are less prevalent. @barr_effect_2020 also addressed this triple interaction between the level of inequality in a society, the individual level of education, and the perceived origin of the disparity (either by luck or effort) to determine the extent to which inequality is accepted. They found that the interaction's mechanism varies depending on the compared societies. Finally, @garcia-sanchez_attitudes_2020, using data for 41 countries from the International Social Survey Programme (ISSP), found that the perceived size of the income gap correlated positively with support for progressive taxation. Still, this association was weaker among those who endorsed meritocratic and equal opportunity beliefs. In the same line, experimental research by @durante_preferences_2009 subjects support less redistribution when the initial distribution is determined according to task performance.

Research about meritocratic beliefs at school age is rather scarce, leaving a wide research gap as schools are one of the primary socialization institutions where achievement based on merit explains success [@erivwo_meritocracy_2021]. Nevertheless, it is possible to find some initial works focused on the area of distributive justice at school that are closely related to meritocratic beliefs, as the ones by Resh and Sabbagh (quotes). Using justice in grade obtention as a measure of distributive justice and meritocracy (Sabbagh et al 2006), they find for instance that a larger sense of distributive justice about grades is associated to higher socio-economic status [@resh_sense_2010], have a positive effect on liberal democratic orientation and on trust in people and in formal institutions [@resh_sense_2014], and tend to refrain from violence and to engage to a greater extent in extra-curricular school activity and community volunteering [@resh_sense_2017].

## Children's judgments of inequality, schools, and family background

Research indicates various socialization practices at families and schools during childhood and adolescence that impact dispositional and behavioral tendencies concerning justifications of inequality in adult life. The differences in economic understanding across age groups are consistent with research on cognitive development [@choudhury_social_2006]. Adolescents with mature socio-cognitive abilities tend to express stronger preferences for fairness than infants and children [@wynn_not_2018]. As children grow older, they become more likely to behave fairly, with their early-emerging strict egalitarianism being replaced by an increasing endorsement of fairness principles and engagement in collaborative activities [@huppert_development_2019; @mcauliffe_developmental_2017]. In these activities, their fairness views consider individual contributions, merits, and circumstances [@almas_fairness_2010; @huppert_development_2019; @sigelman_development_1991]. @engelmann_children_2019 claim that children’s sense of fairness emerges at three years old, and we can observe it in collaborative activities, where they accept inequality if the procedure gives everyone an equal chance. Therefore, children at this age respond to unequal distributions based on interpersonal concerns, as they already demand equal respect. In any case, between 3 and 8 years of age, inequitable and anti-meritorious allocations are evaluated more negatively, but equitable and meritorious allocations are not evaluated more positively [@elenbaas_unfairness_2019].

Some research shows that the social environment in which children develop, such as family and school, is associated with their prosocial behaviors by playing an essential role in the transmission of equity norms [@schunk_fairness_2023; @kosse_prosociality_2020]. In fact, schools contribute to institutionalizing and reproducing inequality by promoting values, norms, practices, and languages familiar to higher-class families because the dominant group’s culture shapes educational institutions [@bourdieu_reproduction_1990]. Middle- and upper-class students are better equipped to face academic challenges and are more familiar with academic expectations [@mikus_children_2020]. Such familiarity represents cultural capital in educational contexts because higher-status students come to school ready to meet these expectations and reap the benefits [@jack_no_2016; @khan_privilege_2011]. Conversely, lower-status children lacking cultural capital must catch up while experiencing inequitable comparisons [@goudeau_hidden_2017]. Additionally, academic achievement is treated as the outcome of dispositional factors (e.g., pupils’ efforts and talents or lack of them) rather than the result of differential access to critical resources. Due to the meritocratic frame schools encourage, both low- and high-status individuals believe that students’ success or failure is not due to their family background but rather to differences in efforts and talents [@darnon_where_2018]. In this sense, we believe that the perception of meritocracy can influence students' judgments about market justice preferences. Furthermore, we believe there is a difference between students' perceptions of meritocracy based on their own experience in school and what they perceive in society at large. Consistently, our first two hypotheses are:

H1a: Students who perceive that there is more meritocracy at school will show larger market justice preferences

H1b: Students who perceive that there is more meritocracy in society will show larger market justice preferences

Family background and family socialization practices also contribute to children’s and adolescent’s market justice preferences. For example, @almas_fairness_2017 found that adolescents from low-socioeconomic-status families are likelier to have an egalitarian fairness view and consider an equal distribution as fair in a situation with unequal merits. The authors speculate that differences in socialization practices across status groups might bring about, to a great extent, the fairness views of children and adolescents because social status seems to interact with these evaluations [e.g., @hvidberg_social_2023].

The classic work of Kohn showed that middle-class parents value the expression of internal states and emotions, such as self-control, curiosity, happiness, and consideration, while working-class parents promote deference, obedience, and conformity to authority [@kohn_social_1963; @kohn_class_1969]. Although parents from all social backgrounds encourage individualism in their children, this shared norm translates into different forms in high and low social classes (1999). @acemoglu_obedience_2021 claimed that the values families impart to their children interact with social mobility. Because obedience is a valuable characteristic for employers, in low-wage and social mobility environments, low-income families impart values of obedience to their children to prevent disadvantaging them in labor markets. On the one hand, children from privileged families are socialized to adopt a clear conception of individualism that highlights their internal states, independence, and idiosyncrasies. In contrast, children from disadvantaged families are socialized to support a more balanced view of individualism that considers personal characteristics as resources to overcome collective impediments on the path to upward mobility [@iacoviello_collectivism_2019]. In this way, we believe that there are differences in the socialization of values according to socioeconomic differences that could influence market justice preferences and, therefore, we propose the following hypothesis:

H2: Students from families of higher social status will show larger market justice preferences.

Recent empirical research has demonstrated that the institutional design of schools, coupled with the meritocratic ideology it fosters, significantly influences children’s and adolescents’ views on inequality and deservingness. For example, @jonsson_institutional_2015 study revealed that higher-status adolescents in Sweden tend to perpetuate social class stereotypes while describing the vocational and academic tracks. Academic track students are depicted as wealthy, intelligent, ambitious, and diligent, while vocational track students are characterized as poor, unambitious, unintelligent, and lackadaisical. These stereotypes help individuals maintain a sense of superiority over others and legitimize the prevailing social hierarchies and economic disparities [@jost_attitudinal_2000]

H3: Students from schools of higher social status will show larger market justice preferences.

H4: Students from schools with higher average levels of academic achievement will show larger market justice preferences.

\[interacciones\]

H5: The perception of meritocracy in school and society will moderate the effect of family social status on market justice preferences.

H6: The perception of meritocracy in school and society will moderate the effect of school status on market justice preferences.

H7: The perception of meritocracy in school and society will moderate the effect of school academic achievement on market justice preferences.

### Summary of hypotheses

```{r}
#| fig-cap: "Summary of hypotheses"
```

##