### 2024-05-15 03-methods.qmd
Primer cambios
Cambios en method.qmd y analisis.qmd

---
author: "Equipo Edumer"
format: pdf
editor: visual
---

## Data, variables and method

### Data

The main data source corresponds to the First Study of Civic Education in Chile, conducted by the Agency for Quality Education of the Ministry of Education in 2017. This database is based on the administration of a civic knowledge test to students and questionnaires to various members of the educational community. The target population consists of 8th-grade students in 242 schools across the country. Initially, the database comprises 8,589 student observations and 6,770 parent observations. However, considering the survey's unit of analysis and to ensure higher data quality, 171 student cases and 79 parent cases exhibiting repetitive and careless response patterns were removed (Gottfried et al., 2022). Additionally, data from the System for Measuring the Quality of Education (SIMCE) conducted by the Ministry of Education for the year 2017 are employed, providing school-level information such as administrative dependence, socioeconomic classification, and results obtained in standardized mathematics and language tests. After variable processing and removal of missing cases, the final database for this study includes a two-level stratified sample, composed of 5,047 students and parents (level 1) nested within 231 schools (level 2).

### Variables

***Individual level***

**Justification of Inequality**: The dependent variable in this study is the justification of social inequality in specific policy domains, namely health, education, and pensions. This construct is measured through three variables addressing the degree of justification regarding whether access to social services in these domains should be income conditional. Justification of inequality in health is measured by the item "Is it fair in Chile that people with higher incomes can access better healthcare than those with lower incomes," while in education, it is measured using the statement "Is it fair in Chile that people who can afford it have better education for their children." In the case of pensions, it is measured through the item "Is it fair in Chile that people with higher incomes can have better pensions than those with lower incomes." In all cases, respondents declare their preferences on a Likert scale ranging from "completely disagree" (1) to "completely agree" (4). Additionally, we include a summarized indicator of preferences for market justice (REF), measured by an average index across these items (α = 0.86), with values ranging from 1 to 4, where higher values represent greater preferences for market justice (see @tbl-desc-dependientes).

```{r echo=FALSE}
pacman::p_load(summarytools, dplyr, sjlabelled)
load("input/data/dependientes.RData")
dependientes$bienestar <- set_label(dependientes$bienestar, "Market Justice Preferences")
```

```{r, results='asis', echo=FALSE}
#| tbl-cap: "Dependent variables"
#| label: tbl-desc-dependientes

df<-dfSummary(dependientes,
               plain.ascii = FALSE,
               style = "multiline",
               tmp.img.dir = "/tmp",
               graph.magnif = 0.75,
               headings = F,  # encabezado
               varnumbers = F, # num variable
               labels.col = T, # etiquetas
               na.col = F,    # missing
               graph.col = F, # plot
               valid.col = T, # n valido
               col.widths = c(30,10,10,10))

df$Variable <- NULL # delete variable column

print(df)
```

**Perception of Meritocracy**: The main independent variable refers to the perception of meritocracy, operationalized through five items addressing the perception of rewards based on talent and intelligence at both the school and societal levels. At the school level, students respond to whether "Intelligence is important for getting good grades" and "Effort is important for getting good grades". At the societal level, students respond to the following questions: "In Chile, people are rewarded for their effort", "In Chile, people get what they deserve" and "In Chile, people are rewarded for their intelligence and skills". Each item was answered on a four-point Likert scale ranging from "completely disagree" (1) to "completely agree" (4).

The analysis includes several individual control variables that have been identified as influential in recent research. These variables encompass family socioeconomic status, assessed by the highest level of parental education, the number of books in the household, and an index measuring access to technologies, all aimed at mitigating compositional effects (REFS). @tbl-desc-independent shows the independent variables used, their response categories and their frequencies.

```{r echo=FALSE}
load("input/data/independientes.RData")
```

```{r, results='asis', echo=FALSE}
#| tbl-cap: "Independent variables"
#| label: tbl-desc-independent

df<-dfSummary(independientes,
               plain.ascii = FALSE,
               style = "multiline",
               tmp.img.dir = "/tmp",
               graph.magnif = 0.75,
               headings = F,  # encabezado
               varnumbers = F, # num variable
               labels.col = T, # etiquetas
               na.col = F,    # missing
               graph.col = F, # plot
               valid.col = T, # n valido
               col.widths = c(30,10,10,10))

df$Variable <- NULL # delete variable column

print(df)
```

***Contextual level***

This study examines four variables at the school level. Firstly, we assess the administrative dependence of schools, categorized as "Public," "Subsidized Private," or "Private." Secondly, we consider the socioeconomic classification of schools, determined by the Ministry of Education and measured on an ordinal scale with five categories ranging from "low" (1) to "high" (5). Thirdly, we evaluate the performance level of schools in SIMCE standardized tests, categorized as "low," "medium," or "high." Lastly, we analyze the proportion of parents with university or postgraduate education level at each school. The items used, response categories, and their frequencies are detailed in @tbl-desc-school.

```{r echo=FALSE}
load("input/data/school.RData")
```

```{r, results='asis', echo=FALSE}
#| tbl-cap: "School context variables"
#| label: tbl-desc-school

df<-dfSummary(school,
               plain.ascii = FALSE,
               style = "multiline",
               tmp.img.dir = "/tmp",
               graph.magnif = 0.75,
               headings = F,  # encabezado
               varnumbers = F, # num variable
               labels.col = T, # etiquetas
               na.col = F,    # missing
               graph.col = F, # plot
               valid.col = T, # n valido
               col.widths = c(30,10,10,10))

df$Variable <- NULL # delete variable column

print(df)
```

### Methods

Given the hierarchical structure of the data, with students nested within schools, model estimation is performed within a multilevel framework. These models are appropriate for capturing both individual and contextual effects in a single analysis, allowing for the estimation of fixed effects between groups and random effects that vary from one group to another [@bell_fixed_2019; @hox_multilevel_2010]. Cumulative link mixed models were employed for the ordinal dependent variables, while linear mixed-effects models were applied for the average market justice preference index.

The hypotheses of this research were pre-registered in the Open Science Framework platform of the Center for Open Science (OSF), the access to the document is available at this [link](https://doi.org/10.17605/OSF.IO/UFSDV). The statistical analysis of this research was performed using the free software R version 4.1.3.