### 2024-05-17 02-antecedentes.qmd
revisada introduccion y antecedentes, temas menores formato

---
author: "Equipo EDUMER"
editor: visual
bibliography: "input/bib/just-ineq-merit.bib"
csl: "input/bib/apa6.csl"
---

## Justification of inequality and market justice

The justification of social inequality based on marke-type criteria has been conceptualized as the individuals’ adherence to the deservingness of social goods and services (sucha as health, education and pensions) based on prices and individuals’ ability to pay [@boltanski_new_2005; @lane_market_1986; @streeck_citizens_2012]. Research on social stratification beliefs, which explore individual perceptions of who deserves what and why [@kluegel_beliefs_1987], highlights that people’s explanations and justifications of social inequality are closely tied to their judgments of deservingness. The influence of ideologies [@wegener_dominant_1995] and cultural schemas [@homan_being_2017] is pivotal in shaping these explanations by offering symbolic representations that frame societal structures and expectations. While significant attention has been paid to wage inequality, income distribution, and payment differentials in the literature [@castillo_legitimacy_2011; Evans et al., 2010; @jasso_how_1999 ; @shariff_income_2016], there has been less examination of public beliefs about which life domains should be governed by market relations and even less about children’s acceptance or rejection of these market principles. This oversight is notable given the extensive encroachment of market logic into public goods, welfare policy, and social services over the past five decades [@centeno_arc_2012; @harvey_breve_2015], affecting areas such as pensions, health services, and education.

Robert E. Lane proposed the underpinnings of the concept of market justice, which he differentiated from political justice. For him, “it is the genius of the market to stimulate wants without at the same time stimulating a sense of deserving more than one gets” [-@lane_market_1986; p. 384]. Contrary to the evidence that unequal distribution produces feelings of dissatisfaction, anger, and resentment that might motivate forms of collective action [@greitemeyer_subjective_2016; @mishra_subjective_2015; @smith_relative_2012; @power_deprivationprotest_2018; @power_relative_2020], Lane points out that in market settings, social comparisons are more likely to motivate increased effort rather than feelings of acute injustice because individuals attribute outcomes to their actions. In this sense, unequal levels of well-being ought to be, to some extent, a function of their talents and efforts, instead of being based in distributive principles that characterize welfarism, such as need and equality (see Wilson 2003).

The management of social services manifests in varying approaches across nations, with substantial differences in funding and delivery methods [@jensen_worlds_2008; @stoy_worlds_2014]. Nordic countries, for example, predominantly employ public agencies to produce and provide social services, funding these through collective taxation and offering them in kind to the majority of citizens. This system prioritizes political justice, placing it above market mechanisms in accessing services. In contrast, other countries rely more heavily on for-profit entities and private funding, where service distribution depends mainly on individual financial capacity of paying user fees, highlighting the influence of market justice in service allocation. The trend toward marketization of welfare services has been growing since the 1980s [@salamon_marketization_1993], and this shift is increasingly evident even in countries where market solutions have traditionally had a minor role in social policy [@sivesind_changing_2017]. The expansion of marketization has been related to larger justification of market mechanisms, whereby societies with larger private spending on services show larger market justice preferences [@lindh_public_2015].

Despite high-income inequality and limited social mobility in CHile and in Latin America in general, there is a prevalent belief that individuals are solely responsible for their economic outcomes, a view that varies across the region [@bucca_merit_2016; @chong_mystery_2008; @torche_intergenerational_2014; @salgado_inequality_2023]. The reliance on private welfare providers and widespread user fees [@molyneux_neoliberal_2008] adds complexity to this context. Yet, research on children’s justification of market-based inequalities in accessing welfare services remains limited, highlighting a significant gap in understanding how younger generations view market-based access to welfare and whether these views are associated with their meritocratic beliefs.

## Meritocratic perceptions and market justice

The original definition of merit is a combination of effort and talent [@young_rise_1958], and a meritocracy is a distributive system where merit is the main criteria for allocating valuable goods and rewards. From a sociological perspective, meritocracy has been used in research on social mobility to characterize societies with low mobility that threaten the meritocratic ideal [@goldthorpe_myth_2003]. More recently, sociology and social psychology research has attended to subjective aspects related to the support for meritocratic principles in different societies, such as beliefs in meritocracy [@castillo_multidimensional_2023a; @mijs_unfulfillable_2016; @mijs_paradox_2019]. The label of beliefs in this realm covers different concepts, but in this paper we refer to a conceptual model developed by @castillo_multidimensional_2023 which identify two main aspects: preferences and perceptions for meritocracy.

While meritocratic preferences refer to a justification of distribution based on merit criteria (effort and talent), the perception of meritocracy refers to how individuals view and understand the concept of meritocracy in their own society [@duru-bellat_who_2012; @castillo_meritocracia_2019]. This perception can vary greatly depending on individual experiences, social, economic, and cultural background. Some people may see meritocracy as a fair and just system that allows anyone to succeed based on their abilities and hard work. In contrast, others may view it as a myth or a cover for existing power dynamics and inequality, serving to maintain and even reinforce inequality [@lampert_meritocratic_2013; @mijs_paradox_2021]. Based in this last perspective, we argue that individuals with a higher perception of meritocracy will show a larger market justice preferences, as individual achievement would be seen as rewarded and social policies as less necessary.

Some studies have analyzed how those with greater privileges believe more in meritocracy (Reynolds & Chan 2014), how greater economic inequality increases meritocratic beliefs [@mijs_paradox_2021], and how larger inequality affects meritocratic beliefs [@morris_representing_2022]. A larger justification of meritocratic distribution have been related to less support for redistributive compensation systems [@frank_performance_2015]. Conversely, individuals tend to support redistribution when they believe that the disadvantaged lack the opportunities to succeed [@evans_strong_2018]. @almas_cutthroat_2020 found that in the US the highly educated accept inequality significantly more than the less educated because they perceive inequality as justifiable owing to differences in productivity (i.e., merit). @barr_effect_2020 found that in unequal societies, the highly educated accept inequality more than the low educated. Using data for 41 countries from the International Social Survey Programme (ISSP),@garcia-sanchez_attitudes_2020 found that the perceived size of the income gap correlated positively with support for progressive taxation, an association that was weaker among those who endorsed meritocratic and equal opportunity beliefs. In the same line, experimental research by @durante_preferences_2009 showed that subjects support less redistribution when the initial distribution is determined according to task performance.

## The socialization of market justice: families and schools

Schools contribute to institutionalizing and reproducing inequality by promoting values, norms, practices, and languages familiar to higher-class families because the dominant group’s culture shapes educational institutions [@bourdieu_reproduction_1990]. Middle- and upper-class students are better equipped to face academic challenges and are more familiar with academic expectations [@mikus_children_2020]. Such familiarity represents cultural capital in educational contexts because higher-status students come to school ready to meet these expectations and reap the benefits [@jack_no_2016; @khan_privilege_2011]. Conversely, lower-status children lacking cultural capital must catch up while experiencing inequitable comparisons [@goudeau_hidden_2017]. Additionally, academic achievement is treated as the outcome of dispositional factors (e.g., pupils’ efforts and talents or lack of them) rather than the result of differential access to critical resources. Due to the meritocratic frame schools encourage, both low- and high-status individuals tend to believe that success or failure is not due to family background but rather to differences in efforts and talents [@darnon_where_2018]. In this sense, we believe that the perception of meritocracy can influence students' judgments about market justice preferences.

H1a: Students who perceive that there is more meritocracy in society will show larger market justice preferences

The perception of meritocracy has been mostly studied with general questions about rewards allocation based on effort and talent (ususally ingelligence). Nevertheless, when looking at school population it is possible to further consider the perception of meritocracy referred specifically to the school context. For instance, @resh_sense_2010 find that perception of justice in grades have a positive effect on liberal democratic orientation and on trust in people and in formal institutions [@resh_sense_2014]. Attending to this evidence, we differentiate in this study between meritocratic perception in society at large, and meritocratic perception at school, proposing the next hypothesis:

H1b: Students who perceive that there is more meritocracy at school will show larger market justice preferences

Attending to the socialization within the family, the classic work of Kohn showed that middle-class parents value the expression of internal states and emotions, such as self-control, curiosity, happiness, and consideration, while working-class parents promote deference, obedience, and conformity to authority [@kohn_social_1963; @kohn_class_1969]. Although parents from all social backgrounds encourage individualism in their children, this shared norm translates into different forms in high and low social classes (1999). @acemoglu_obedience_2021 claimed that the values families impart to their children interact with social mobility. On the one hand, children from privileged families are socialized to adopt a clear conception of individualism that highlights their internal states, independence, and idiosyncrasies. In contrast, children from disadvantaged families are socialized to support a more balanced view of individualism that considers personal characteristics as resources to overcome collective impediments on the path to upward mobility [@iacoviello_collectivism_2019]. In the same line, @almas_fairness_2017 found that adolescents from low-socioeconomic-status families are likelier to have an egalitarian fairness view and consider an equal distribution as fair in a situation with unequal merits. Taking this into account, we believe that there will be differences in the socialization of values according to socioeconomic differences in families that could influence market justice preferences, which lead us to the following hypothesis:

H2: Students from families of higher social status will show larger market justice preferences.

Recent empirical research has demonstrated that the institutional design of schools, coupled with the meritocratic ideology it fosters, significantly influences children’s and adolescents’ views on inequality and deservingness. For example, @jonsson_institutional_2015 study revealed that higher-status adolescents in Sweden tend to perpetuate social class stereotypes while describing the vocational and academic tracks. Academic track students are depicted as wealthy, intelligent, ambitious, and diligent, while vocational track students are characterized as poor, unambitious, unintelligent, and lackadaisical. These stereotypes help individuals maintain a sense of superiority over others and legitimize the prevailing social hierarchies and economic disparities [@jost_attitudinal_2000]. The link between school socioeconomic status and market preferences is certainly reinforced in societies with high school segregation, as is the case of Chile. Consistently, we expect that

H3: Students from schools of higher social status will show larger market justice preferences.

The last set of hypotheses deal with the role of meritocratic perception as a moderator of family and school status on market justice preferences. These hypotheses have more an exploratory character, as there are not clear references in the literature to this regard. The general argument deals with the role of socioeconomic status and meritocratic perceptions, which is contested in the literature. On the one hand, and based on a mere rational argument, it would be expected that those who succeed in terms of educational rewards and carreer paths - usuaylly those of higher status - perceive more meritocracy. Nevertheless, evidence regarding the legitimacy of meritocracy relies on consensual beliefs from the side of both winners and losers from the distributive system, pointing in the direction of fewer status differences when it comes to meritocratic beliefs [@wiederkehr_belief_2015]. As this is the first study of this kind in the Chilean context, charaterized by extreme educational inequalities, we rely on a traditional and rational approach, proposing that:

H4: Students from schools of higher social status will perceive more meritocracy.

H5: Students from families of higher social status will perceive more meritocracy.

[aca falta hipothesis de school performance]

The previous hypotheses serve as a sort of bridge for the final moderation hypotheses. As higher (family and school) status would be related to meritocratic perception (H5 and H6), and at the same time meritocratic perception would be related to larger market justice preferences (H1 and H2), we argue that meritocratic perceptions could play a moderating role: the association between high status and market justice would be stronger for those with a larger meritocratic perception:

H6: The perception of meritocracy in school and society will moderate the effect of family social status on market justice preferences.

H7: The perception of meritocracy in school and society will moderate the effect of school status on market justice preferences.

### Summary of hypotheses @Kevin-carrasco, actualizar esquema

```{r echo=FALSE, out.width = "600px"}
#| fig-cap: "Summary of hypotheses"
#| label: fig-hypotheses

knitr::include_graphics("input/img/hypotheses.png",
  auto_pdf = getOption("knitr.graphics.auto_pdf", TRUE))
```