### 2024-05-17 05-conclusiones.qmd

---
author: "Equipo EDUMER"
editor: visual
---

# Discussion

# Conclusions

-   Efecto ilustrador de la educación? (esfuerzo -\> redistribución)