### 2024-06-10 05-conclusiones.qmd
discussion

---
author: "Equipo EDUMER"
editor: visual
---

# Discussion

Our research aimed to examine the relationship between the perception of meritocracy and the justification of market justice preferences among eighth-grade students in Chile. Overall, our findings provided mixed evidence regarding our hypotheses. While we found support for some associations, others did not align with our initial expectations. Hypothesis H1a posited that students who perceive greater meritocracy in society would show stronger preferences for market justice. The results confirm this hypothesis, indicating a positive relationship between the perception that talent and effort are rewarded in society and the preferences for market justice in access to services such as health, education, and pensions. This finding aligns with previous research suggesting that individuals who believe in a meritocratic system are more likely to justify social inequalities [@batruch_belief_2022; @wiederkehr_belief_2015]. Hypothesis H1b asserted that students who perceive greater meritocracy in school would show stronger preferences for market justice. The results show a mixed effect. The perception that talent is important for achieving good grades in school is associated with a greater preference for market justice in access to social services. Still, the perception that effort is important is associated with less preference for market justice. This highlights the influence of the school environment in shaping attitudes toward market justice, suggesting that meritocratic experiences in school can translate to broader societal views [@resh_sense_2010; @resh_sense_2014].

Hypothesis H2 proposed that students from higher socioeconomic status families would show stronger preferences for market justice. However, the results indicate that this relationship is not as straightforward as expected. Although students from families with higher educational levels and more economic resources tend to prefer less market justice, this relationship is not statistically significant. This might be due to the influence of other contextual and cultural factors that modulate attitudes toward market justice [@kohn_social_1963; @kohn_class_1969; @almas_fairness_2017].

Hypothesis H3 held that students from schools with higher socioeconomic status would show stronger preferences for market justice. The results did not confirm this hypothesis since although students from high-status schools tend to prefer more market justice, this result is not statistically significant. This finding, contrary to the existing literature [@jonsson_institutional_2015; @jost_attitudinal_2000], does not allow us to affirm the role of the school environment, particularly socioeconomic segregation, in influencing attitudes towards market justice and the acceptance of inequalities.

Hypothesis Hx held that students from schools with higher school achievement would show stronger preferences for market justice. Contrary to that hypothesis, the results show that students from schools with greater achievement in the SIMCE test tend to prefer less market justice.

Hypotheses H6 and H7 proposed that the perception of meritocracy would moderate the effect of socioeconomic status (both family and school) on market justice preferences. The results show that the perception of meritocracy, especially social meritocracy, amplifies the market justice preferences in schools with low levels of achievement and in families with parents with a lower educational level. This means that although those who perceive society as meritocratic have greater preferences towards market justice, this influence is less among those who have a higher socioeconomic status and attend schools with higher academic achievement. This finding highlights the complex interplay between socioeconomic status, school context, meritocratic beliefs, and preferences towards market justice.

# Conclusions

Future research could leverage the capabilities of longitudinal databases to explore contextual level. For example, statistical models could incorporate contextual information from administrative data sources, allowing for testing hypotheses that include contextual socioeconomic inequality and its dynamics over time. Additionally, this dataset could be used to compare inequality justification across different policy areas, such as health and pensions, and examine how meritocratic perceptions and beliefs affect these areas differently over time. Such studies would provide valuable insights for understanding societal and cultural changes and their impacts on solidarity and social cohesion.

-   Efecto ilustrador de la educación? (esfuerzo -\> redistribución)