### 2024-06-14 01-introduction.qmd
revisión párrafos chile

---
author: "Equipo EDUMER"
format: html
bibliography: "input/bib/just-ineq-merit.bib"
csl: "input/bib/apa6.csl"
---

# Introduction

Since its origins, educational institutions have been related to the idea of social mobility and access to better opportunities. Therefore, the consistent evidence of the high level of social reproduction at the school level represents a threat to the promise of education and a meritocratic system [@bourdieu_reproduction_1990]. A large part of the research in this field at an international level has addressed the extent to which the social origin of students affects their academic results and their life opportunities [@vonhippel_test_2019], confirming that schools have severe difficulties in closing the socio-economic and cultural gaps of origin. Besides this socioeconomic perspective on school opportunities, recent research has addressed to what extent inequalities in the school context are also influencing students' perceptions, beliefs, and attitudes: Are social inequalities even perceived at the school context? Are they rejected by the students, particularly those who are worst-off in socioeconomic terms? Or, Is there evidence at the school level that social inequalities are tolerated and even justified? [@batruch_belief_2022; @wiederkehr_belief_2015].

Given that the school environment has an important focus on performance, achievement and acknowledgment, meritocracy has been one of the principal concepts used for understanding and even for justifying performance differences among students. Meritocracy is a distributive system based on the belief that people should be rewarded and promoted based on their abilities, knowledge, and achievements [@young_rise_1958]. It is often seen as a way to create equal opportunities and fairness, as individuals can rise to positions of power and influence based on their own merit rather than their background or connections. However, some argue that meritocracy can actually lead to tolerating or even justifying social inequalities, as it can create a hierarchy where those who already have resources and advantages are more likely to succeed. In this regard, a great deal of academic research about meritocracy delves into the assessment of to what extent rewards and privileges in society are related to merit, emphasizing the so-called unfulfillable promise of meritocracy [@mijs_stratified_2016].

In the present paper, we address the role of the perception of meritocracy on the justification of social inequalities by eighth-grade students in Chile, a country characterized by a highly stratified educational system. In particular, we focus in justifying inequalities in health, education and pensions, which are traditional social policy areas where acces to better services could be justified based on payment capacity, referred as to market justice [@lane_market_1986, @lindh_bringing_2023].\
Most of the research in this realm has only considered adults, leaving aside the study of how beliefs in this field develop at student age as well as the impact of the school context and the family as the main socialization agencies [@batruch_belief_2022]. Regarding schools, the way in which they deal with unequal conditions of origin has been linked to the *hidden curriculum* [@chafel_schooling_1997], whereby students learn about distributive norms in society and mechanisms of justification of social differences.

Based on recent studies that relate school meritocracy to the justification of economic inequalities in the adult population [@batruch_belief_2022; @wiederkehr_belief_2015], the central hypothesis guiding this research is that school-age students with a higher perception of meritocracy - both at school and at the societal level - will show a larger justification of social inequalities, as individual achievement would be seen as appropriately rewarded and social mechanisms for correcting inequalities as less necessary [@batruch_belief_2022]. We focus on the student-age population as we point out that it is possible to track down the origin of meritocratic beliefs (and their consequences) to early socialization processes. To this regard, we take into account the family and the school as two main socialization agencies that play a significant role in the socialization of cultural beliefs by transmitting cultural norms, values, and expectations to young people.

The Chilean case is particularly intriguing for studying market justice preferences. This country is characterized by acute and persistent economic inequality, which stands out in Latin America and among OECD countries [@flores_top_2020]. In Chile, the poorest 50% captures only 10% of the total income and has negative wealth, while the richest 1% receives almost 27% of the income and holds 49.6% of the wealth [@chancel_world_2022]. Much of this inequality has been attributed to the deep neoliberal reforms that institutionalized privatization and commodification of various economic sectors. Such reforms where introduced during the dictatorship (1973-1989) and expanded in democracy through concessions, demand credits, and specific regulatory frameworks [@ffrench-davis_reformas_2018]. This shift in economic policy has allowed the unprecedented emergence of markets in areas related to social reproduction, such as health, pensions, and education, with provision and access managed by private entities and segmented by individual payment capacity, heavily reliant on State subsidies [@boccardo_30_2020]. In health, although the majority of the population uses the public insurance system (78.9%), the pension system is based on individual capitalization, with mandatory contributions managed by private administrators investing in the financial market, currently involving 11 million contributors (Superintendencia de Pensiones, 2024). In education, 30.6% of school enrollment is in public schools, 54.0% in state-subsidized (voucher) private schools, and 9.3% in fully private schools, generally attracting higher-income groups (Mineduc, 2023).

Some surveys in Chile have considered the perception and support for the market provision in public areas. Recent public opinion surveys conducted by the Center for Public Studies (CEP) indicate that, generally, people prefer public insurance (64.1%) over private insurance (35.9%) if both cost the same. In pensions, Chileans show varied preferences regarding the public or private status of pension fund administrators; 52.3% favor both private and public institutions, 38.8% prefer only public institutions, and only 8.6% prefer only private ones. This likely relates to the severe legitimacy crisis faced by private pension fund administrators, which has sparked social movements in recent years [@boccardo_30_2020]. Interestingly, in education, there is a clear preference for private schools over public ones if both cost the same and are equally distant, with 63% preferring private and 37% public.