### 2024-06-18 paper.qmd

---
format:
  html: 
    toc: true
    keep-tex: true
smooth-scroll: true
number-sections: true
bibliography: "input/bib/just-ineq-merit.bib"
csl: "input/bib/apa6.csl"
link-citations: true
mainfont: Times New Roman
---

# The socialization of meritocracy and market justice preferences at school {.unnumbered}

Abstract

{{< include 01-introduction.qmd >}}

{{< include 02-antecedentes.qmd >}}

{{< include 03-methods.qmd >}}

{{< include 04-analisis.qmd >}}

{{< include 05-discussion.qmd >}}

{{< include 06-conclusion.qmd >}}

{{< include 07-anexos.qmd >}}