### 2024-06-21 processing/issp2019.Rmd

---
title: "ISSP2019"
author: "Equipo EDUMER"
date: "2022-11-28"
output: html_document
---

```{r}
pacman::p_load(dplyr, ggplot2, sjPlot, haven, sjlabelled, tidyverse)

data <- read_sav("input/data/original/ISSP2019.sav")
```

```{r}
table(data$v30)
summary(data$v30)
data$v30 <- factor(data$v30, 
                   levels = c(5,4,3,2,1),
                   labels = c("Very unjust, definitely wrong", 
                              "Somewhat unjust, wrong", 
                              "Neither just nor unjust, mixed feelings", 
                              "Somewhat just, right", 
                              "Very just, definitely right"
                              ))

data$v30 <- set_label(data$v30, label = "Just or unjust that people with higher incomes can buy better health care than people with lower incomes?")

table(data$v31)
summary(data$v31)
data$v31 <- factor(data$v31, 
                   levels = c(5,4,3,2,1),
                   labels = c("Very unjust, definitely wrong", 
                              "Somewhat unjust, wrong", 
                              "Neither just nor unjust, mixed feelings", 
                              "Somewhat just, right", 
                              "Very just, definitely right"
                              ))

data$v31 <- set_label(data$v31, label = "Just or unjust that people with higher incomes can buy better education for children than people with lower incomes?")
```

```{r}
issp2019 <- sjPlot::plot_stackfrq(dplyr::select(data, v30, v31), geom.size = 0.4) +
  theme_bw() +
  theme(legend.position="bottom")
issp2019

ggsave(issp2019, file = "output/graphs/issp2019.png",device = "png",width = 25,height = 13,dpi = "retina",units = "cm")
```

# Cleveland plot

```{r}
data$v30 <- as.numeric(data$v30)
data$v31 <- as.numeric(data$v31)

data$country <- factor(data$country, labels = get_labels(data$country))

health <- data %>% group_by(country) %>% 
  summarise(Health = mean(v30, na.rm=TRUE)) %>% 
  arrange(Health) %>% 
  mutate(country = factor(country, levels = .$country))

education <- data %>% group_by(country) %>%
  summarise(Education = mean(v31, na.rm=TRUE)) %>% 
  arrange(Education) %>% 
  mutate(country = factor(country, levels = .$country))

base <- left_join(health, education, by="country")

base <- base[order(base$Health), ]

base <- base %>% pivot_longer(c("Health", "Education"), 
                              names_to="justice", 
                              values_to="mean")
```

```{r}
cleveland_issp <- base %>% ggplot(aes(mean, country)) +
  theme_bw() +
  geom_point(aes(color = justice, shape = justice), size = 3) +
  stat_summary(aes(xintercept = after_stat(x), y = 0), fun = mean, geom = "vline", orientation = "y", colour = "green") +
  xlab("average") +
  scale_x_continuous(limits = c(1, 5),
                     breaks = seq(1, 5, by = 1),
                     label = c("Very unjust, definitely wrong", 
                               "Somewhat unjust, wrong", 
                               "Neither just nor unjust, 
  mixed feelings", 
                               "Somewhat just, right", 
                               "Very just, definitely right"))
cleveland_issp

ggsave(cleveland_issp, file = "output/graphs/cleveland_issp.png",device = "png",width = 25,height = 13,dpi = "retina",units = "cm")
```

# Solo Chile

```{r}
data_chile <- data %>% filter(country==152)
data_chile$v30 <- as.numeric(data_chile$v30)
data_chile$v31 <- as.numeric(data_chile$v31)
issp2019_chile <- sjPlot::plot_stackfrq(select(data_chile, v30, v31), geom.size = 0.4) +
  theme_bw() +
  theme(legend.position="bottom")
issp2019_chile

ggsave(issp2019_chile, file = "output/graphs/issp2019_chile.png",device = "png",width = 25,height = 13,dpi = "retina",units = "cm")
```