### 2024-06-24 06-conclusion.qmd

---
author: "Equipo EDUMER"
bibliography: "input/bib/just-ineq-merit.bib"
csl: "input/bib/apa6.csl"
---

# Conclusion

Our research aimed to examine the relationship between the perception of meritocracy and the justification of market justice preferences among eighth-grade students in Chile. From our knowledge this is the first study addressing market justice preferences at school level, finding that students exhibit large preferences for market justice (about one third agree or strongly agree with them), a share that is quite larger when compared with evidence in adult population. Perception of meritocracy is also high, particularly when it comes to the reward of effort, with an striking eighty six percent of agreement. Regarding these descriptive indicators, a a first message coming from this study is that there is something particular in school-age when it comes to perceptions and preferences about inequality. This can relate to development aspects (Kim & Gewirtz, 2019, Rizzo & Killen, 2020) but it could be something specific to the Chilean case, which requires further comparative as well as longitudinal research in the area. In any case, teaching about the origins and characteristics of social inequality is not part of the chilean school curriculum, calling for a more dedicated consideration of these topics in areas such as citizensip education as well as in history and social science education.

The core of this paper was the association between meritocratic perceptions and preferences for market justice. In general, we find that those who perceive that effort and talent/intelligence are rewarded, are more willing to agree with that richer individuals can have better health, education, and pensions. This is not the first time that meritocratic perceptions (and also beliefs) have been related with the legitimation of social inequalities (Mijs, 2019; Darnon 2017), but the consideration of school level population as well as market justice preferences allows adding evidence regarding the role of meritocracy regarding inequality beliefs: legitimation of market criteria in social services seems to begin at school age, and what students think of meritocracy is linked to it. Although there are still several open questions, as for instance differences between meritocratic perceptions in school and in the society at large, this raises up the relevance of what is done (or not) in schools in order to challenge the meritocratic ideals.

The consideration of the school context opened some caveats in this research. We found contrary evidence for the association between status, school performance, and market justice prefecences: high status and high achievement at school level is asociated with less market justice preferences. This is opposed to the rational-interest argument that suggests that those doing better - both economically and academically- would demand less redistribution (Gonthier,2016; Runst, 2018; Bullock, 2021) and therefore are expected to challenge market mechanisms. A possible explanation is that education contributes to form a crirical view of markets in redistribution, nevertheless further reserach in this area is needed in order to assert the specific mechanisms that might be playing a role. All in all, from this study we know that school contexts are not trivial to this regard, calling for larger attention to school age research on the formation of inequality preferences and beliefs.