### 2024-06-28 05-discussion.qmd
revision JC

---
author: "Equipo EDUMER"
editor: visual
---

# Discussion

The previous analyses aimed to explore the relationship between perceptions of meritocracy and preferences for market justice among students in Chilean society. Several models estimated how students' beliefs about the rewards for talent and effort in both societal and school contexts influence their attitudes toward access to social services like health, education, and pensions based on pay capacity. Additionally, the models examined the impact of students' socioeconomic status and school achievement on these preferences, assessing whether perceptions of meritocracy moderate these effects. The findings offer insights into the complex interplay between meritocratic beliefs, socioeconomic factors, and attitudes toward market justice criteria.

Hypothesis $H_{1a}$ posited that students who perceive greater meritocracy in society would show stronger preferences for market justice. The results support this hypothesis, indicating a positive relationship between the perception that talent and effort are rewarded in society and the preferences for market justice in access to services such as health, education, and pensions. This finding aligns with previous research suggesting that individuals who believe in a meritocratic system are more likely to justify social inequalities [@batruch_belief_2022; @wiederkehr_belief_2015]. It is noteworthy that among the items measuring meritocratic perception, consistently the one about deservingness shows a larger effect when compared to the ones about talent and effort. Although the three items correlate, there is something distinctive about deservingness that probably encompasses other aspects beyond meritocracy, an aspect to which we will come back in the conclusions. Moving now to hypothesis $H_{1b}$, this is simply a variation of 1a, replacing the perception of meritocracy at society by the one at school. In this case we have only two items of meritocratic perception - unfortunately deservingness was not measured at school level - and the direction of the effects is not as straightforward as in the case of societal meritocracy in $H_{1a}$: positive for talent, but negative for effort. Still, one of the limitation for the comparison between perception of meritocracy and society is the difference of the items' phrasing. Whereas for society the item refers to the extent to to which efforts are rewarded, at school is about the importance of effort, whereby something that is important does not necessarily mean that it is enough to be rewarded. Furthermore, the school-effort item is the one with the lowest variation among all meritocratic items (more than 95% agree or strongly agree), and with the lowest effect size.

Hypothesis $H_2$ and hypothesis $H_3$ propose that socioeconomic status could influence students' preferences, but the results show different patterns. On the one hand, $H_2$ proposed that students from higher socioeconomic status families would show stronger preferences for market justice, which is not supported by the results. Although students from families with higher educational levels and more economic resources tend to prefer less market justice, this relationship is not statistically significant. This might be due to the influence of other contextual and cultural factors that modulate attitudes toward market justice [@kohn_social_1963; @kohn_class_1969; @almas_fairness_2017]. Notably, socioeconomic status measured by parental educational level and the number of books at home do not explain variations in market justice preferences, suggesting that other variables, such as market justice preferences of parents, could play an essential role in the attitudinal socialization process. On the other hand, Hypothesis $H_3$ held that students from schools with higher socioeconomic status would show stronger preferences for market justice. The results confirm this hypothesis since students from high-status schools tend to prefer more market justice, compared to low-status schools. This finding, in line with the existing literature [@jonsson_institutional_2015; @jost_attitudinal_2000], does allow us to affirm the hypothesis about the role of the school environment, particularly socioeconomic segregation, in influencing attitudes towards market justice and the acceptance of inequalities. A possible explanation is the role of the context at the school level, which could have a more relevant role than the family socioeconomic status, since the highly stratified nature of the Chilean education system reinforces socioeconomic divisions and socializes students into different normative frameworks.

Hypothesis $H_4$ held that students from schools with higher school achievement would show stronger preferences for market justice, following the argument that such schools could foster meritocratic attitudes by showing a larger promotion of talent and effort in order to get better results. Nevertheless, the results show quite the opposite: schools with better academic results (as measured by a national standardized test) show in average less preferences for market justice. This finding is one of the most strongest and consistent throughout the models and it is relevant given the centrality of school performance in educational institutions. One possible explanation of this link is that students in high-achieving schools may be more exposed to discussions and a deeper understanding of inequality and social justice, leading to a critical perspective toward the weight of market mechanisms in the access to social services.

Hypotheses $H_5$ and $H_6$ proposed that socioeconomic status would moderate the effect of meritocratic perceptions on market justice preferences. Hypothesis $H_5$ refers to family socioeconomic status, whereby the results show a negative interaction with two of the three perceptual variables at the societal level. This means that the positive impact of the perception of social meritocracy on market justice preferences had a lower effect on students with parents with higher education and also on those with more books at home. Something similar occurs with hypothesis $H_6$, where now school-level status moderates the effect of meritocracy variables, which in multilevel framework is referred as to cross-level interaction. The negative interaction in this context means that in schools of low socioeconomic levels, the positive effect of the social meritocracy variables on market justice preferences is stronger. These relations indicate that, although those who perceive society as meritocratic have greater preferences towards market justice, this influence is lower among those with a higher socioeconomic level. This finding highlights the complex interaction between socioeconomic status, school context, meritocratic beliefs, and preferences toward market justice. The moderating role of meritocratic perceptions suggests that students' views on social justice are shaped by their socioeconomic background and beliefs about how society rewards effort and talent. This underlines the importance of addressing structural inequalities and individual beliefs in educational and social policies.

Hypothesis $H7$ posited that the academic performance of schools could moderate the effect of meritocratic perceptions on preferences for market justice. This hypothesis is based on the argument that the level of academic performance in schools entails a differentiated promotion of meritocratic perceptions to achieve better results, which would, in turn, reinforce or diminish preferences for market justice. The results of the interactions between levels suggest that, indeed, in higher-performing schools, the effect of meritocratic perceptions on the justification of service provision based on market principles is weaker compared to lower-performing schools. Similarly to the effect of academic performance in $H4$, a possible explanation for this moderation is that students in high-performing schools may be more exposed to knowledge about the effects of inequality and social justice. This exposure translates into an enlightening effect of education on inequality, which can mitigate the impact of meritocracy on market justice preferences.