### 2024-07-04 07-supplementary-material.qmd

---
author: "Equipo EDUMER"
bibliography: "input/bib/just-ineq-merit.bib"
csl: "input/bib/apa6.csl"
---

## Supplementary Material {#appendix}

```{r}
#| tbl-cap: "Translation of inequality justification items"
```

|                                                                                                                    |                                                                                                                                              |
|---------------------------------|---------------------------------------|
| **English translation**                                                                                            | **Original Spanish**                                                                                                                         |
| It is just that in Chile people who cna pay have a better education for their children                             | Es justo que en Chile las personas que puedan pagar tengan una mejor educación para sus hijos.                                               |
| It is just that in Chile people with higher incomes can have better pensions than people with low incomes          | Es justo que en Chile las personas con mayores ingresos puedan tener mejores pensiones que las personas de ingresos más bajos                |
| It is just that in Chile people with higher incomes can access better health services than people with low incomes | Es justo que en Chile las personas con mayores ingresos puedan acceder a una mejor atención de salud que las personas con ingresos más bajos |

```{r}
#| fig-cap: "Access to better health services"
```

```{r}
#| fig-cap: "Access to better education"
```

```{r}
#| fig-cap: "Access to better pensions"
```