### 2024-07-04 back_matter.qmd

---
title: "The Socialization of Meritocracy and Market Justice Preferences at School"
subtitle: "Back Matter"
format:
  hikmah-pdf: 
    indent: false
  hikmah-manuscript-pdf: 
    indent: false
bibliography: "input/bib/just-ineq-merit.bib"
csl: "input/bib/apa6.csl"
link-citations: true
linestretch: 1.15       
mainfont: Times New Roman
fontsize: 12pt          
geometry: "left=2.54cm,right=2.54cm,top=2.54cm,bottom=2.54cm"
papersize: letter
colorlinks: true
fig-height: 4
fig-width: 7.5
header-includes:
  - \usepackage{threeparttable}
---

{{< include 07-supplementary-material.qmd >}}

## Acknowledgments

The authors acknowledge the support of ANID/FONDECYT grant 1210847 and the Center for Conflict and Social Cohesion Studies (COES) grant ANID/FONDAP 1513009.

## Authors' Contributions

The following statements should be used "Conceptualization, X.X. and Y.Y.; Methodology, X.X.; Software, X.X.; Validation, X.X., Y.Y. and Z.Z.; Formal Analysis, X.X.; Investigation, X.X.; Resources, X.X.; Data Curation, X.X.; Writing – Original Draft Preparation, X.X.; Writing – Review & Editing, X.X.; Visualization, X.X.; Supervision, X.X.; Project Administration, X.X.; Funding Acquisition, Y.Y.”

## Conflict of Interest

The authors report no competing interests to declare.

## Funding

This research received no external funding.

## Data Avaibility Statement

The original data presented in the study are openly available in <https://www.agenciaeducacion.cl/orientar/estudios/estudios-nacionales/>

Access the data, code, and analysis presented in the study are openly available in <https://github.com/educacion-meritocracia/market-justice-preferences-school>.

## References