### 2024-07-04 paper.qmd

---
title: "The socialization of meritocracy and market justice preferences at school"
format:
  hikmah-pdf: 
    indent: false
  hikmah-manuscript-pdf: 
    indent: false
author:
  - name: Juan Carlos Castillo
    affiliations: Department of Sociology, Universidad de Chile
    email: juancastillov@uchile.cl
    attributes:
      corresponding: true
  - name: Mauricio Salgado
    affiliations: School of Sociology, Universidad Andrés Bello
  - name: Kevin Carrasco
    affiliations: Center for Social Conflict and Cohesion Studies
  - name: Andreas Laffert
    affiliations: Institute of Sociology, Universidad Católica de Chile
abstract: |
  This article evaluates novel approaches to do some really important things.
keywords: 
  - open-source 
  - scientific publishing
  - reproducible research
bibliography: "input/bib/just-ineq-merit.bib"
csl: "input/bib/apa6.csl"
link-citations: true
linestretch: 1.15       
mainfont: Times New Roman
fontsize: 12pt          
geometry: "left=2.54cm,right=2.54cm,top=2.54cm,bottom=2.54cm"
papersize: letter
colorlinks: true
fig-height: 4
fig-width: 7.5
header-includes:
  - \usepackage{threeparttable}
---

{{< include 01-introduction.qmd >}}

{{< include 02-antecedentes.qmd >}}

{{< include 03-methods.qmd >}}

{{< include 04-analisis.qmd >}}

{{< include 05-discussion.qmd >}}

{{< include 06-conclusion.qmd >}}