### 2024-07-08 07-supplementary-material.qmd
Ultimos cambios: paper, front matter, cover letter, back matter

---
author: "Equipo EDUMER"
bibliography: "input/bib/just-ineq-merit.bib"
csl: "input/bib/multidisciplinary-digital-publishing-institute.csl"
---

# Supplementary Material {#appendix}

```{r load, include=FALSE}
pacman::p_load(tidyverse,
               ggpubr, 
               lme4,
               ordinal,
               sjPlot, 
               texreg,
               kableExtra,
               easystats,
               ggeffects,
               knitr,
               webshot2)

options(scipen=999)
rm(list = ls())

load("input/data/data_index.RData")
load("input/data/data_rec.RData")
load("input/data/mlm_models.RData")
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F,
                      warning = F,
                      error = F, 
                      message = F) 
```

## Translated items for market justice preferences

```{r echo=FALSE}
#| tbl-cap: "Translation of items for health, pensions, and education"
#| label: tbl-translation

data <- data.frame(
  English = c("It is just that in Chile people with higher incomes can access better health services than people with low incomes",
              "It is just that in Chile people with higher incomes can have better pensions than people with low incomes",
              "It is just that in Chile people who can pay have a better education for their children"),
  Spanish = c("Es justo que en Chile las personas con mayores ingresos puedan acceder a una mejor atención de salud que las personas con ingresos más bajos",
              "Es justo que en Chile las personas con mayores ingresos puedan tener mejores pensiones que las personas de ingresos más bajos",
              "Es justo que en Chile las personas que puedan pagar tengan una mejor educación para sus hijos")
  )

kableExtra::kable(data, 
      col.names = c("English Translation", "Original Spanish"), 
      booktabs = T, escape = F, align = 'l') %>%
  kableExtra::kable_styling(latex_options = "HOLD_position", bootstrap_options = "striped", full_width = T)  %>%
  row_spec(0, bold = T)

```

## Complete multilevel models

```{r echo=FALSE}

data_rec <- data_rec %>% mutate(universitario = case_when(educacion_rec=="Ed Basica"~"No",
                                                          educacion_rec=="Ed Media"~"No",
                                                          educacion_rec=="Ed Tecnica"~"No",
                                                          educacion_rec=="Universidad o posgrado"~"Si",
                                                          educacion_rec=="Ns/Nr"~"Ns/Nr"))
data_rec$universitario <- factor(data_rec$universitario, levels = c("No", "Si", "Ns/Nr"))

data_rec <- data_rec %>% mutate(school_nse = case_when(cod_grupo=="Bajo"~"Bajo",
                                                          cod_grupo=="Medio bajo"~"Medio",
                                                          cod_grupo=="Medio"~"Medio",
                                                          cod_grupo=="Medio alto"~"Medio",
                                                          cod_grupo=="Alto"~"Alto"))
data_rec$school_nse <- factor(data_rec$school_nse, levels = c("Bajo", "Medio", "Alto"))

data_salud <- data_rec %>% 
    dplyr::select(-just_pension, -just_educ, -redistribucion)
data_salud <- na.omit(data_salud)
data_salud$just_salud_fact <- as_factor(data_salud$just_salud)

m0_h <- clmm(just_salud_fact ~ 1 + (1 | mrbd), data=data_salud)

m1_h <- clmm(just_salud_fact ~ 1 + inteligencia_esc + esfuerzo_esc + 
    inteligencia_soc + esfuerzo_soc + merito_soc + (1 | mrbd), data=data_salud)

m2_h <- clmm(just_salud_fact ~ 1 + inteligencia_esc + esfuerzo_esc + 
    inteligencia_soc + esfuerzo_soc + merito_soc + universitario + 
    libros_rec + (1 | mrbd), data=data_salud)

m3_h <- clmm(just_salud_fact ~ 1 + inteligencia_esc + esfuerzo_esc + 
    inteligencia_soc + esfuerzo_soc + merito_soc + universitario + 
    libros_rec + acc_tec + internet_rec + (1 | mrbd), data=data_salud)

m4_h <- clmm(just_salud_fact ~ 1 + inteligencia_esc + esfuerzo_esc + 
    inteligencia_soc + esfuerzo_soc + merito_soc + universitario + 
    libros_rec + acc_tec + internet_rec + school_nse + simce + (1 | mrbd), data=data_salud)

m5_h <- clmm(just_salud_fact ~ 1 + inteligencia_esc + esfuerzo_esc + 
    inteligencia_soc + esfuerzo_soc + merito_soc + universitario + 
    libros_rec + acc_tec + internet_rec + school_nse + simce +
    mean_educ + cod_depe2 +(1 | mrbd), data=data_salud)

modelos_health <- list(m0_h, m1_h, m2_h, m3_h, m4_h, m5_h)

```

```{r echo=FALSE, results='asis'}
#| tbl-cap: "Cumulative Link Multilevel Models for Justifying Differential Access to Health Care"
#| label: tbl-ordinal-health

ccoef <- list(
  "1|2" = "Strongly Disagree|Disagree",
  "2|3" = "Disagree|Agree",
  "3|4" = "Agree|Strongly Agree",
  inteligencia_esc = "School talent",
  esfuerzo_esc = "School effort",
  inteligencia_soc = "Social talent",
  esfuerzo_soc = "Social effort",
  merito_soc = "Deservingness",
  universitarioSi = "University or posgraduate",
  "universitarioNs/Nr" = "Missing",
  "libros_recMás de 25" = "More than 25 books (Ref.= Less than 25)",
  acc_tec = "Technology acces",
  internet_rec2 = "No internet connectivity (Ref.= Internet connectivity)",
  school_nseMedio = "SES Medium",
  school_nseAlto = "SES High",
  simceMedio = "Simce Medium",
  simceAlto = "Simce High",
  mean_educ = "Prop. university level at school",
  "cod_depe2Part. subvencionado" = "Subsidized Private",
  "cod_depe2Part. privado" = "Private")

texreg::texreg(
  l = modelos_health,
  caption = NULL,
  stars = c(0.05, 0.01, 0.001),
  custom.note = "\\item Note: Cells contain regression coefficients with standard errors in parentheses. %stars \\\\ \\item Source: own elaboration based on First Study of Civic Education 2017.",
  threeparttable = T,
  custom.model.names = paste0(rep("Model ",6), seq(0,5)),
  custom.coef.map = ccoef,
  leading.zero = TRUE,
  use.packages = FALSE,
  booktabs = TRUE,
  scalebox = 0.8,
  groups = list("Parental education (Ref.= Non-university)" = 9:10, "Socioeconomic level (Ref.= Low)" = 14:15, "Achievement score (Ref.= Low)" = 16:17, "School dependence (Ref.= Public)" = 19:20),
  custom.gof.names = c("Log Likelihood",
                       "AIC", 
                       "BIC", 
                       "Num.obs",
                       "Num. groups: Schools",
                       "Var: Schools (Intercept)"))

```

```{r echo=FALSE}
data_pension <- data_rec %>% 
    dplyr::select(-just_educ, -just_salud, -redistribucion)
data_pension <- na.omit(data_pension)
data_pension$just_pension_fact <- as_factor(data_pension$just_pension)

m0_p <- clmm(just_pension_fact ~ 1 + (1 | mrbd), data=data_pension)

m1_p <- clmm(just_pension_fact ~ 1 + inteligencia_esc +
             esfuerzo_esc + inteligencia_soc + esfuerzo_soc +
             merito_soc + (1 | mrbd), data=data_pension)

m2_p <- clmm(just_pension_fact ~ 1 + inteligencia_esc +
             esfuerzo_esc + inteligencia_soc + esfuerzo_soc +
             merito_soc + universitario + libros_rec +
             (1 | mrbd), data=data_pension)

m3_p <- clmm(just_pension_fact ~ 1 + inteligencia_esc +
             esfuerzo_esc + inteligencia_soc + esfuerzo_soc +
             merito_soc + universitario + libros_rec + acc_tec +
             internet_rec + (1 | mrbd), data=data_pension)

m4_p <- clmm(just_pension_fact ~ 1 + inteligencia_esc +
             esfuerzo_esc + inteligencia_soc + esfuerzo_soc +
             merito_soc + universitario + libros_rec + acc_tec + 
             internet_rec + school_nse + simce + (1 | mrbd),
             data=data_pension)

m5_p <- clmm(just_pension_fact ~ 1 + inteligencia_esc +
             esfuerzo_esc + inteligencia_soc + esfuerzo_soc +
             merito_soc + universitario + libros_rec + acc_tec + 
             internet_rec + school_nse + simce + mean_educ + 
             cod_depe2 + (1 | mrbd), data=data_pension)

modelos_pension <- list(m0_p, m1_p, m2_p, m3_p, m4_p, m5_p)

```

```{r echo=FALSE, results='asis'}
#| tbl-cap: "Cumulative Link Multilevel Models for Justifying Differential Access to Pensions"
#| label: tbl-ordinal-pension

texreg::texreg(
  l = modelos_pension,
  caption = NULL,
  stars = c(0.05, 0.01, 0.001),
  custom.note = "\\item Note: Cells contain regression coefficients with standard errors in parentheses. %stars \\\\ \\item Source: own elaboration based on First Study of Civic Education 2017.",
  threeparttable = T,
  custom.model.names = paste0(rep("Model ",6), seq(0,5)),
  custom.coef.map = ccoef,
  leading.zero = TRUE,
  use.packages = FALSE,
  booktabs = TRUE,
  scalebox = 0.7,
  groups = list("Parental education (Ref.= Non-university)" = 9:10, "Socioeconomic level (Ref.= Low)" = 14:15, "Achievement score (Ref.= Low)" = 16:17, "School dependence (Ref.= Public)" = 19:20),
  custom.gof.names = c("Log Likelihood",
                       "AIC", 
                       "BIC", 
                       "Num.obs",
                       "Num. groups: Schools",
                       "Var: Schools (Intercept)"))

```

```{r echo=FALSE}
data_educ <- data_rec %>% 
    dplyr::select(-just_pension, -just_salud, -redistribucion)
data_educ <- na.omit(data_educ)
data_educ$just_educ_fact <- as_factor(data_educ$just_educ)

m0_e <- clmm(just_educ_fact ~ 1 + (1 | mrbd), data=data_educ)

m1_e <- clmm(just_educ_fact ~ 1 + inteligencia_esc + esfuerzo_esc + 
    inteligencia_soc + esfuerzo_soc + merito_soc + (1 | mrbd), data=data_educ)

m2_e <- clmm(just_educ_fact ~ 1 + inteligencia_esc + esfuerzo_esc + 
    inteligencia_soc + esfuerzo_soc + merito_soc + universitario + 
    libros_rec + (1 | mrbd), data=data_educ)

m3_e <- clmm(just_educ_fact ~ 1 + inteligencia_esc + esfuerzo_esc + 
    inteligencia_soc + esfuerzo_soc + merito_soc + universitario + 
    libros_rec + acc_tec + internet_rec + (1 | mrbd), data=data_educ)

m4_e <- clmm(just_educ_fact ~ 1 + inteligencia_esc + esfuerzo_esc + 
    inteligencia_soc + esfuerzo_soc + merito_soc + universitario + 
    libros_rec + acc_tec + internet_rec + school_nse + simce + (1 | mrbd), data=data_educ)

m5_e <- clmm(just_educ_fact ~ 1 + inteligencia_esc + esfuerzo_esc + 
    inteligencia_soc + esfuerzo_soc + merito_soc + universitario + 
    libros_rec + acc_tec + internet_rec + school_nse + simce +
    mean_educ + cod_depe2 + (1 | mrbd), data=data_educ)

modelos_educ <- list(m0_e, m1_e, m2_e, m3_e, m4_e, m5_e)

```

```{r echo=FALSE, results='asis'}
#| tbl-cap: "Cumulative Link Multilevel Models for Justifying Differential Access to Education"
#| label: tbl-ordinal-educ

texreg::texreg(
  l = modelos_educ,
  caption = NULL,
  stars = c(0.05, 0.01, 0.001),
  custom.note = "\\item Note: Cells contain regression coefficients with standard errors in parentheses. %stars \\\\ \\item Source: own elaboration based on First Study of Civic Education 2017.",
  threeparttable = T,
  custom.model.names = paste0(rep("Model ",6), seq(0,5)),
  custom.coef.map = ccoef,
  leading.zero = TRUE,
  use.packages = FALSE,
  booktabs = TRUE,
  scalebox = 0.7,
  groups = list("Parental education (Ref.= Non-university)" = 9:10, "Socioeconomic level (Ref.= Low)" = 14:15, "Achievement score (Ref.= Low)" = 16:17, "School dependence (Ref.= Public)" = 19:20),
  custom.gof.names = c("Log Likelihood",
                       "AIC", 
                       "BIC", 
                       "Num.obs",
                       "Num. groups: Schools",
                       "Var: Schools (Intercept)"))

```

```{r echo=FALSE, include=FALSE}

data_bienestar <- data_bienestar %>% mutate(universitario = case_when(educacion_rec=="Ed Basica"~"No",
                                                          educacion_rec=="Ed Media"~"No",
                                                          educacion_rec=="Ed Tecnica"~"No",
                                                          educacion_rec=="Universidad o posgrado"~"Si",
                                                          educacion_rec=="Ns/Nr"~"Ns/Nr"))
data_bienestar$universitario <- factor(data_bienestar$universitario, levels = c("No", "Si", "Ns/Nr"))

data_bienestar <- data_bienestar %>% mutate(school_nse = case_when(cod_grupo=="Bajo"~"Bajo",
                                                          cod_grupo=="Medio bajo"~"Medio",
                                                          cod_grupo=="Medio"~"Medio",
                                                          cod_grupo=="Medio alto"~"Medio",
                                                          cod_grupo=="Alto"~"Alto"))
data_bienestar$school_nse <- factor(data_bienestar$school_nse, levels = c("Bajo", "Medio", "Alto"))

reg0_bienestar <- lmer(bienestar ~ 1 + (1 | mrbd), data = data_bienestar)

reg1_bienestar <- lmer(bienestar ~ 1 + inteligencia_esc + esfuerzo_esc + 
                         inteligencia_soc + esfuerzo_soc + merito_soc + 
                         (1 | mrbd), data=data_bienestar)

reg2_bienestar <- lmer(bienestar ~ 1 + inteligencia_esc + esfuerzo_esc + 
                         inteligencia_soc + esfuerzo_soc + merito_soc + 
                         universitario + libros_rec + (1 | mrbd), data=data_bienestar)

reg3_bienestar <- lmer(bienestar ~ 1 + inteligencia_esc + esfuerzo_esc + 
                         inteligencia_soc + esfuerzo_soc + merito_soc + 
                         universitario + libros_rec + acc_tec + 
                         internet_rec +  (1 | mrbd), data=data_bienestar)

reg4_bienestar <- lmer(bienestar ~ 1 + inteligencia_esc + esfuerzo_esc + 
                         inteligencia_soc + esfuerzo_soc + merito_soc + 
                         universitario + libros_rec + acc_tec + 
                         internet_rec +  school_nse + simce + 
                         (1 | mrbd), data=data_bienestar)

reg5_bienestar <- lmer(bienestar ~ 1 + inteligencia_esc + esfuerzo_esc + 
                         inteligencia_soc + esfuerzo_soc + merito_soc + 
                         universitario + libros_rec + acc_tec + 
                         internet_rec +  school_nse + simce + 
                         mean_educ + cod_depe2 + 
                         (1 | mrbd), data=data_bienestar)
```

```{r echo=FALSE, results='asis'}
#| tbl-cap: "Linear mixed-effects models for meritocracy perception and market justice preferences"
#| label: tbl-lineal-reg
ccoef <- list(
 "(Intercept)" = "Constant",
  inteligencia_esc = "School talent",
  esfuerzo_esc = "School effort",
  inteligencia_soc = "Social talent",
  esfuerzo_soc = "Social effort",
  merito_soc = "Deservingness",
  universitarioSi = "University or posgraduate",
  "universitarioNs/Nr" = "Missing",
  "libros_recMás de 25" = "More than 25 books (Ref.= Less than 25)",
  acc_tec = "Technology acces",
  internet_rec2 = "No internet connectivity (Ref.= Internet connectivity)",
  school_nseMedio = "SES Medium",
  school_nseAlto = "SES High",
  simceMedio = "Simce Medium",
  simceAlto = "Simce High",
  mean_educ = "Prop. university level at school",
  "cod_depe2Part. subvencionado" = "Subsidized Private",
  "cod_depe2Part. privado" = "Private")

texreg::texreg(
  l = list(reg0_bienestar, reg1_bienestar, reg2_bienestar, reg3_bienestar, reg4_bienestar, reg5_bienestar),
  caption = NULL,
  stars = c(0.05, 0.01, 0.001),
  custom.note = "\\item Note: Cells contain regression coefficients with standard errors in parentheses. %stars \\\\ \\item Source: own elaboration based on First Study of Civic Education 2017.",
  threeparttable = T,
  custom.model.names = paste0(rep("Model ",6), seq(0,5)),
  custom.coef.map = ccoef,
  leading.zero = TRUE,
  use.packages = FALSE,
  booktabs = TRUE,
  scalebox = 0.7,
  groups = list("Parental education (Ref.= Non-university)" = 7:8, "Socioeconomic level (Ref.= Low)" = 12:13, "Achievement score (Ref.= Low)" = 14:15, "School dependence (Ref.= Public)" = 16:18),
   custom.gof.names = c("AIC", 
                               "BIC", 
                               "Log Likelihood", 
                               "Num.obs",
                               "Num. groups: Schools",
                               "Var: Schools (Intercept)",
                               "Var: Residual"))

```

```{r}

load("input/data/data_rec.RData")

data_rec <- data_rec %>% rowwise() %>%  mutate(educacion_rec2 = case_when(educacion_rec=="Ed Basica"~"Low level",
                                                 educacion_rec=="Ed Media"~"Low level",
                                                 educacion_rec=="Ed Tecnica"~"Low level",
                                                 educacion_rec=="Universidad o posgrado"~"University or Postgraduate",
                                                 educacion_rec=="Ns/Nr"~"Missing"
                                                 ))
data_rec$educacion_rec2 <- factor(data_rec$educacion_rec2, levels = c("Low level", "University or Postgraduate", "Missing"))

data_rec <- data_rec %>% rowwise() %>%  mutate(cod_grupo2  = case_when(cod_grupo=="Bajo"~"Bajo",
                                                                       cod_grupo=="Medio bajo"~"Bajo",
                                                                       cod_grupo=="Medio"~"Bajo",
                                                                       cod_grupo=="Medio alto"~"Alto",
                                                                       cod_grupo=="Alto"~"Alto"))
data_rec$cod_grupo2 <- factor(data_rec$cod_grupo2, levels = c("Bajo", "Alto"))
data_rec$simce <- factor(data_rec$simce, labels = c("Low", "Medium", "High"))

data_bienestar <- na.omit(data_rec)
#data_bienestar <- data_bienestar %>% mutate(pension_rec = 5-just_pension,
#                                            educ_rec = 5-just_educ,
#                                            salud_rec = 5-just_salud)
data_bienestar <- data_bienestar %>% rowwise() %>% dplyr::mutate(bienestar = mean(c(just_educ, just_salud, just_pension), na.omit=TRUE))
data_bienestar$bienestar <- round(data_bienestar$bienestar, 2)

```

```{r}

# meritocracia x ses familiar (educacion)
reg10_bienestar <- lmer(bienestar ~ 1 + merito_soc + inteligencia_soc + esfuerzo_esc + inteligencia_esc + educacion_rec2*esfuerzo_soc + libros_rec + acc_tec + mean_educ + cod_depe2 + cod_grupo2 + simce + internet_rec + (1 + educacion_rec2 | mrbd), data=data_bienestar)

reg11_bienestar <- lmer(bienestar ~ 1 + esfuerzo_soc + inteligencia_soc + esfuerzo_esc + inteligencia_esc + educacion_rec2*merito_soc + libros_rec + acc_tec + mean_educ + cod_depe2 + cod_grupo2 + simce + internet_rec + (1 + educacion_rec2 | mrbd), data=data_bienestar)

reg12_bienestar <- lmer(bienestar ~ 1 + esfuerzo_soc + merito_soc + esfuerzo_esc + inteligencia_esc + educacion_rec2*inteligencia_soc + libros_rec + acc_tec + mean_educ + cod_depe2 + cod_grupo2 + simce + internet_rec + (1 + educacion_rec2 | mrbd), data=data_bienestar)

reg13_bienestar <- lmer(bienestar ~ 1 + esfuerzo_soc + merito_soc + inteligencia_soc + inteligencia_esc + educacion_rec2*esfuerzo_esc + libros_rec + acc_tec + mean_educ + cod_depe2 + cod_grupo2 + simce + internet_rec + (1 + educacion_rec2 | mrbd), data=data_bienestar)

reg14_bienestar <- lmer(bienestar ~ 1 + esfuerzo_soc + merito_soc + inteligencia_soc + esfuerzo_esc + educacion_rec2*inteligencia_esc + libros_rec + acc_tec + mean_educ + cod_depe2 + cod_grupo2 + simce + internet_rec + (1 + educacion_rec2 | mrbd), data=data_bienestar)
```

```{r echo=FALSE, results='asis'}
#| tbl-cap: "Linear mixed-effects models for meritocracy perception, parental education and market justice preferences"
#| label: tbl-interac-educparent

ccoef <- list(
  "(Intercept)" = "Constant",
  inteligencia_esc = "School talent",
  esfuerzo_esc = "School effort",
  inteligencia_soc = "Social talent",
  esfuerzo_soc = "Social effort",
  merito_soc = "Deservingness",
  "educacion_rec2University or Postgraduate" = "University or posgraduate",
  "educacion_rec2Missing" = "Missing",
  "libros_recMás de 25" = "More than 25 books (Ref.= Less than 25)",
  cod_grupo2Alto = "SES High (Ref.= Low)",
  simceMedium = "Simce Medium",
  simceHigh = "Simce High",
  "educacion_rec2University or Postgraduate:esfuerzo_soc" = "University or posgraduate * Social effort",
  "educacion_rec2Missing:esfuerzo_soc" = "Missing * Social effort",
  "educacion_rec2University or Postgraduate:merito_soc" = "University or posgraduate * Deservingness",
  "educacion_rec2Missing:merito_soc" = "Missing * Deservingness",
  "educacion_rec2University or Postgraduate:inteligencia_soc" = "University or posgraduate * Social talent",
  "educacion_rec2Missing:inteligencia_soc" = "Missing * Social talent",
  "educacion_rec2University or Postgraduate:esfuerzo_esc" = "University or posgraduate * School effort",
  "educacion_rec2Missing:esfuerzo_esc" = "Missing * School effort",
  "educacion_rec2University or Postgraduate:inteligencia_esc" = "University or posgraduate * School talent",
  "educacion_rec2Missing:inteligencia_esc" = "Missing * School talent"
  )

# meritocracia x ses familiar (educacion)

texreg::texreg(l = list(reg10_bienestar, reg11_bienestar,
                   reg12_bienestar, reg13_bienestar,
                   reg14_bienestar),
          caption = NULL,
          stars = c(0.05, 0.01, 0.001),
          custom.note = "\\item Note: Cells contain regression coefficients with standard errors in parentheses. Control variables are included. %stars \\\\ \\item Source: own elaboration based on First Study of Civic Education 2017.",
          threeparttable = T,
          custom.model.names = paste0(rep("Model ",5), seq(1,5)),
          custom.coef.map = ccoef,
          leading.zero = TRUE,
          use.packages = FALSE,
          booktabs = TRUE,
          scalebox = 0.7,
          groups = list("Parental education (Ref.= Non-university)" = 7:8,
                        "Achievement score (Ref.= Low)" = 11:12,
                        "Parental education x Meritocracy" = 13:22),
          custom.gof.names = c("AIC", 
                               "BIC", 
                               "Log Likelihood", 
                               "Num.obs",
                               "Num. groups: Schools",
                               "Var: Schools (Intercept)",
                               "Var: Schools University or Postgraduate",
                               "Var: Schools Missing",
                               "Cov: Schools (Intercept) University or Postgraduate",
                               "Cov: Schools (Intercept) Missing",
                               "Cov: Schools University or Postgraduate, Missing",
                               "Var: Residual"),
          custom.gof.rows = list("Controls"=rep("Yes",5)))

```

```{r}

# meritocracia x ses familiar (libros)

reg15_bienestar <- lmer(bienestar ~ 1 + merito_soc + inteligencia_soc + esfuerzo_esc + inteligencia_esc + educacion_rec2 + libros_rec*esfuerzo_soc + acc_tec + mean_educ + cod_depe2 + cod_grupo2 + simce + internet_rec + (1 + libros_rec | mrbd), data=data_bienestar)
reg16_bienestar <- lmer(bienestar ~ 1 + esfuerzo_soc + inteligencia_soc + esfuerzo_esc + inteligencia_esc + educacion_rec2 + libros_rec*merito_soc + acc_tec + mean_educ + cod_depe2 + cod_grupo2 + simce + internet_rec + (1 + libros_rec | mrbd), data=data_bienestar)
reg17_bienestar <- lmer(bienestar ~ 1 + esfuerzo_soc + merito_soc + esfuerzo_esc + inteligencia_esc + educacion_rec2 + libros_rec*inteligencia_soc + acc_tec + mean_educ + cod_depe2 + cod_grupo2 + simce + internet_rec + (1 + libros_rec | mrbd), data=data_bienestar)
reg18_bienestar <- lmer(bienestar ~ 1 + esfuerzo_soc + merito_soc + inteligencia_soc + inteligencia_esc + educacion_rec2 + libros_rec*esfuerzo_esc + acc_tec + mean_educ + cod_depe2 + cod_grupo2 + simce + internet_rec + (1 + libros_rec | mrbd), data=data_bienestar)
reg19_bienestar <- lmer(bienestar ~ 1 + esfuerzo_soc + merito_soc + inteligencia_soc + esfuerzo_esc + educacion_rec2 + libros_rec*inteligencia_esc + acc_tec + mean_educ + cod_depe2 + cod_grupo2 + simce + internet_rec + (1 + libros_rec | mrbd), data=data_bienestar)
```

```{r echo=FALSE, results='asis'}
#| tbl-cap: "Linear mixed-effects models for meritocracy perception, books at home and market justice preferences"
#| label: tbl-interac-books

ccoef <- list(
  "(Intercept)" = "Constant",
  inteligencia_esc = "School talent",
  esfuerzo_esc = "School effort",
  inteligencia_soc = "Social talent",
  esfuerzo_soc = "Social effort",
  merito_soc = "Deservingness",
  "educacion_rec2University or Postgraduate" = "University or posgraduate",
  "educacion_rec2Missing" = "Missing",
  "libros_recMás de 25" = "More than 25 books (Ref.= Less than 25)",
  cod_grupo2Alto = "SES High (Ref.= Low)",
  simceMedium = "Simce Medium",
  simceHigh = "Simce High",
  "libros_recMás de 25:esfuerzo_soc" = "More than 25 books * Social effort",
  "libros_recMás de 25:merito_soc" = "More than 25 books * Deservingness",
  "libros_recMás de 25:inteligencia_soc" = "More than 25 books * Social talent",
  "libros_recMás de 25:esfuerzo_esc" = "More than 25 books * School effort",
  "libros_recMás de 25:inteligencia_esc" = "More than 25 books * School talent"
  )

# meritocracia x ses familiar (libros)

texreg::texreg(l = list(reg15_bienestar, reg16_bienestar,
                   reg17_bienestar, reg18_bienestar,
                   reg19_bienestar),
          caption = NULL,
          stars = c(0.05, 0.01, 0.001),
          custom.note = "\\item Note: Cells contain regression coefficients with standard errors in parentheses. Control variables are included. %stars \\\\ \\item Source: own elaboration based on First Study of Civic Education 2017.",
          threeparttable = T,
          custom.model.names = paste0(rep("Model ",5), seq(1,5)),
          custom.coef.map = ccoef,
          leading.zero = TRUE,
          use.packages = FALSE,
          booktabs = TRUE,
          scalebox = 0.7,
          groups = list("Parental education (Ref.= Non-university)" = 7:8,
                        "Achievement score (Ref.= Low)" = 11:12,
                        "Books at home x Meritocracy" = 13:17),
          custom.gof.names = c("AIC", 
                               "BIC", 
                               "Log Likelihood", 
                               "Num.obs",
                               "Num. groups: Schools",
                               "Var: Schools (Intercept)",
                               "Var: Schools More than 25 books",
                               "Cov: Schools (Intercept) More than 25 books",
                               "Var: Residual"),
          custom.gof.rows = list("Controls"=rep("Yes",5)))

```

```{r}

# meritocracia x ses escuela

reg20_bienestar <- lmer(bienestar ~ 1 + merito_soc + inteligencia_soc + esfuerzo_esc + inteligencia_esc + educacion_rec2 + libros_rec + acc_tec + mean_educ + cod_depe2 + cod_grupo2*esfuerzo_soc + simce + internet_rec + (1 + cod_grupo2 | mrbd), data=data_bienestar)
reg21_bienestar <- lmer(bienestar ~ 1 + esfuerzo_soc + inteligencia_soc + esfuerzo_esc + inteligencia_esc + educacion_rec2 + libros_rec + acc_tec + mean_educ + cod_depe2 + cod_grupo2*merito_soc + simce + internet_rec + (1 + cod_grupo2 | mrbd), data=data_bienestar)
reg22_bienestar <- lmer(bienestar ~ 1 + esfuerzo_soc + merito_soc + esfuerzo_esc + inteligencia_esc + educacion_rec2 + libros_rec + acc_tec + mean_educ + cod_depe2 + cod_grupo2*inteligencia_soc + simce + internet_rec + (1 + cod_grupo2 | mrbd), data=data_bienestar)
reg23_bienestar <- lmer(bienestar ~ 1 + esfuerzo_soc + merito_soc + inteligencia_soc + inteligencia_esc + educacion_rec2 + libros_rec + acc_tec + mean_educ + cod_depe2 + cod_grupo2*esfuerzo_esc + simce + internet_rec + (1 + cod_grupo2 | mrbd), data=data_bienestar)
reg24_bienestar <- lmer(bienestar ~ 1 + esfuerzo_soc + merito_soc + inteligencia_soc + esfuerzo_esc + educacion_rec2 + libros_rec + acc_tec + mean_educ + cod_depe2 + cod_grupo2*inteligencia_esc + simce + internet_rec + (1 + cod_grupo2 | mrbd), data=data_bienestar)
```

```{r echo=FALSE, results='asis'}
#| tbl-cap: "Linear mixed-effects models for meritocracy perception, school socioeconomic status and market justice preferences"
#| label: tbl-interac-schoolses

ccoef <- list(
  "(Intercept)" = "Constant",
  inteligencia_esc = "School talent",
  esfuerzo_esc = "School effort",
  inteligencia_soc = "Social talent",
  esfuerzo_soc = "Social effort",
  merito_soc = "Deservingness",
  "educacion_rec2University or Postgraduate" = "University or posgraduate",
  "educacion_rec2Missing" = "Missing",
  "libros_recMás de 25" = "More than 25 books (Ref.= Less than 25)",
  cod_grupo2Alto = "SES High (Ref.= Low)",
  simceMedium = "Simce Medium",
  simceHigh = "Simce High",
  "cod_grupo2Alto:esfuerzo_soc" = "SES High * Social effort",
  "cod_grupo2Alto:merito_soc" = "SES High * Deservingness",
  "cod_grupo2Alto:inteligencia_soc" = "SES High * Social talent",
  "cod_grupo2Alto:esfuerzo_esc" = "SES High * School effort",
  "cod_grupo2Alto:inteligencia_esc" = "SES High * School talent"
  )

# meritocracia x ses escuela

texreg::texreg(l = list(reg20_bienestar, reg21_bienestar,
                   reg22_bienestar, reg23_bienestar,
                   reg24_bienestar),
          caption = NULL,
          stars = c(0.05, 0.01, 0.001),
          custom.note = "\\item Note: Cells contain regression coefficients with standard errors in parentheses. Control variables are included. %stars \\\\ \\item Source: own elaboration based on First Study of Civic Education 2017.",
          threeparttable = T,
          custom.model.names = paste0(rep("Model ",5), seq(1,5)),
          custom.coef.map = ccoef,
          leading.zero = TRUE,
          use.packages = FALSE,
          booktabs = TRUE,
          scalebox = 0.7,
          groups = list("Parental education (Ref.= Non-university)" = 7:8,
                        "Achievement score (Ref.= Low)" = 11:12,
                        "School SES x Meritocracy" = 13:17),
          custom.gof.names = c("AIC", 
                               "BIC", 
                               "Log Likelihood", 
                               "Num.obs",
                               "Num. groups: Schools",
                               "Var: Schools (Intercept)",
                               "Var: Schools SES High",
                               "Cov: Schools (Intercept) SES High",
                               "Var: Residual"),
          custom.gof.rows = list("Controls"=rep("Yes",5)))

```

```{r}

#meritocracia x prop. educacion universitaria por escuela

reg20.1_bienestar <- lmer(bienestar ~ 1 + merito_soc + inteligencia_soc + esfuerzo_esc + inteligencia_esc + educacion_rec2 + libros_rec + acc_tec + mean_educ*esfuerzo_soc + cod_depe2 + cod_grupo2 + simce + internet_rec + (1 + mean_educ | mrbd), data=data_bienestar)
reg21.1_bienestar <- lmer(bienestar ~ 1 + esfuerzo_soc + inteligencia_soc + esfuerzo_esc + inteligencia_esc + educacion_rec2 + libros_rec + acc_tec + mean_educ*merito_soc + cod_depe2 + cod_grupo2 + simce + internet_rec + (1 + mean_educ | mrbd), data=data_bienestar)
reg22.1_bienestar <- lmer(bienestar ~ 1 + esfuerzo_soc + merito_soc + esfuerzo_esc + inteligencia_esc + educacion_rec2 + libros_rec + acc_tec + mean_educ*inteligencia_soc + cod_depe2 + cod_grupo2 + simce + internet_rec + (1 + mean_educ | mrbd), data=data_bienestar)
reg23.1_bienestar <- lmer(bienestar ~ 1 + esfuerzo_soc + merito_soc + inteligencia_soc + inteligencia_esc + educacion_rec2 + libros_rec + acc_tec + mean_educ*esfuerzo_esc + cod_depe2 + cod_grupo2 + simce + internet_rec + (1 + mean_educ | mrbd), data=data_bienestar)
reg24.1_bienestar <- lmer(bienestar ~ 1 + esfuerzo_soc + merito_soc + inteligencia_soc + esfuerzo_esc + educacion_rec2 + libros_rec + acc_tec + mean_educ*inteligencia_esc + cod_depe2 + cod_grupo2 + simce + internet_rec + (1 + mean_educ | mrbd), data=data_bienestar)
```

```{r}
# meritocracia x simce

reg25_bienestar <- lmer(bienestar ~ 1 + merito_soc + inteligencia_soc + esfuerzo_esc + inteligencia_esc + educacion_rec2 + libros_rec + acc_tec + mean_educ + cod_depe2 + cod_grupo2 + simce*esfuerzo_soc + internet_rec + (1 + simce | mrbd), data=data_bienestar)

reg26_bienestar <- lmer(bienestar ~ 1 + esfuerzo_soc + inteligencia_soc + esfuerzo_esc + inteligencia_esc + educacion_rec2 + libros_rec + acc_tec + mean_educ + cod_depe2 + cod_grupo2 + simce*merito_soc + internet_rec + (1 + simce | mrbd), data=data_bienestar)

reg27_bienestar <- lmer(bienestar ~ 1 + esfuerzo_soc + merito_soc + esfuerzo_esc + inteligencia_esc + educacion_rec2 + libros_rec + acc_tec + mean_educ + cod_depe2 + cod_grupo2 + simce*inteligencia_soc + internet_rec + (1 + simce | mrbd), data=data_bienestar)

reg28_bienestar <- lmer(bienestar ~ 1 + esfuerzo_soc + merito_soc + inteligencia_soc + inteligencia_esc + educacion_rec2 + libros_rec + acc_tec + mean_educ + cod_depe2 + cod_grupo2 + simce*esfuerzo_esc + internet_rec + (1 + simce | mrbd), data=data_bienestar)

reg29_bienestar <- lmer(bienestar ~ 1 + esfuerzo_soc + merito_soc + inteligencia_soc + esfuerzo_esc + educacion_rec2 + libros_rec + acc_tec + mean_educ + cod_depe2 + cod_grupo2 + simce*inteligencia_esc + internet_rec + (1 + simce | mrbd), data=data_bienestar)
```

```{r echo=FALSE, results='asis'}
#| tbl-cap: "Linear mixed-effects models for meritocracy perception, school perfomance status and market justice preferences"
#| label: tbl-interac-schoolperfo

ccoef <- list(
  "(Intercept)" = "Constant",
  inteligencia_esc = "School talent",
  esfuerzo_esc = "School effort",
  inteligencia_soc = "Social talent",
  esfuerzo_soc = "Social effort",
  merito_soc = "Deservingness",
  "educacion_rec2University or Postgraduate" = "University or posgraduate",
  "educacion_rec2Missing" = "Missing",
  "libros_recMás de 25" = "More than 25 books (Ref.= Less than 25)",
  cod_grupo2Alto = "SES High (Ref.= Low)",
  simceMedium = "Simce Medium",
  simceHigh = "Simce High",
  "simceMedium:esfuerzo_soc" = "Simce Medium * Social effort",
  "simceHigh:esfuerzo_soc" = "Simce High * Social effort",
  "simceMedium:merito_soc" = "Simce Medium * Deservingness",
  "simceHigh:merito_soc" = "Simce High * Deservingness",
  "simceMedium:inteligencia_soc" = "Simce Medium * Social talent",
  "simceHigh:inteligencia_soc" = "Simce High * Social talent",
  "simceMedium:esfuerzo_esc" = "Simce Medium * School effort",
  "simceHigh:esfuerzo_esc" = "Simce High * School effort",
  "simceMedium:inteligencia_esc" = "Simce Medium * School talent",
  "simceHigh:inteligencia_esc" = "Simce High * School talent"
  )

# meritocracia x perfomance escuela

texreg::texreg(l = list(reg25_bienestar, reg26_bienestar,
                   reg27_bienestar, reg28_bienestar,
                   reg29_bienestar),
          caption = NULL,
          stars = c(0.05, 0.01, 0.001),
          custom.note = "\\item Note: Cells contain regression coefficients with standard errors in parentheses. Control variables are included. %stars \\\\ \\item Source: own elaboration based on First Study of Civic Education 2017.",
          threeparttable = T,
          custom.model.names = paste0(rep("Model ",5), seq(1,5)),
          custom.coef.map = ccoef,
          leading.zero = TRUE,
          use.packages = FALSE,
          booktabs = TRUE,
          scalebox = 0.7,
          groups = list("Parental education (Ref.= Non-university)" = 7:8,
                        "Achievement score (Ref.= Low)" = 11:12,
                        "School perfomance (SIMCE) x Meritocracy" = 13:22),
          custom.gof.names = c("AIC", 
                               "BIC", 
                               "Log Likelihood", 
                               "Num.obs",
                               "Num. groups: Schools",
                               "Var: Schools (Intercept)",
                               "Var: Schools Simce Medium",
                               "Var: Schools Simce High",
                               "Cov: Schools (Intercept) Simce Medium",
                               "Cov: Schools (Intercept) Simce High",
                               "Cov: Schools Simce Medium, Simce High",
                               "Var: Residual"),
          custom.gof.rows = list("Controls"=rep("Yes",5)))

```