### 2024-07-08 cover_letter.qmd

---
title: "The Socialization of Meritocracy and Market Justice Preferences at School"
subtitle: "Cover Letter"
format:
  hikmah-pdf: 
    indent: false
  hikmah-manuscript-pdf: 
    indent: false
bibliography: "input/bib/just-ineq-merit.bib"
csl: "input/bib/multidisciplinary-digital-publishing-institute.csl"
link-citations: true
linestretch: 1.15       
mainfont: Times New Roman
fontsize: 12pt          
geometry: "left=2.54cm,right=2.54cm,top=2.54cm,bottom=2.54cm"
papersize: letter
colorlinks: true
fig-height: 4
fig-width: 7.5
header-includes:
  - \usepackage{threeparttable}
---

This article evaluates

## Authors Identification

Juan Carlos Castillo, Department of Sociology, Universidad de Chile, Santiago de Chile[^*].  
ORCID: 0000-0003-1265-7854

Mauricio Salgado, School of Sociology, Universidad Andrés Bello.  
ORCID: 0000-0002-8108-6154

Kevin Carrasco, Center for Social Conflict and Cohesion Studies.  
ORCID: 0000-0002-9348-8361

Andreas Laffert, Institute of Sociology, Pontificia Universidad Católica de Chile.  
ORCID: 0000-0002-9008-2454

## Exclusive Submission Statement

The authors confirm that this article has been submitted exclusively to this journal.

[^*]: Corresponding author (juancastillov@uchile.cl - [jc-castillo.com](https://juancarloscastillo.github.io/jc-castillo/) - Ignacio Carrera Pinto 1045 Ñuñoa, Santiago de Chile). The authors acknowledge the support of ANID/FONDECYT grant 1210847 and the Center for Conflict and Social Cohesion Studies (COES) grant ANID/FONDAP 1513009.