### 2024-07-08 paper.qmd

---
title: "The socialization of meritocracy and market justice preferences at school"
format:
  hikmah-pdf: 
    indent: false
  hikmah-manuscript-pdf: 
    indent: false
author:
  - name: Juan Carlos Castillo
    email: juancastillov@uchile.cl
    orcid: 0000-0003-1265-7854
    affiliations: 
        name: Department of Sociology, Universidad de Chile
        country: Chile
    attributes:
      corresponding: true
  - name: Mauricio Salgado
    email: msalgado@unab.cl
    orcid: 0000-0002-8108-6154
    affiliations: 
        name: Centro de Estudios Públicos, Chile
        country: Chile
  - name: Kevin Carrasco
    email: kevin.carrasco@ug.uchile.cl
    orcid: 0000-0002-9348-8361
    affiliations: 
        name: Center for Social Conflict and Cohesion Studies
        country: Chile
  - name: Andreas Laffert
    email: alaffertt@estudiante.uc.cl
    orcid: 0000-0002-9008-2454
    affiliations: 
        name: Institute of Sociology, Universidad Católica de Chile
        country: Chile
abstract: |
  This article evaluates novel approaches to do some really important things.
keywords: 
  - open-source 
  - scientific publishing
  - reproducible research
bibliography: "input/bib/just-ineq-merit.bib"
csl: "input/bib/multidisciplinary-digital-publishing-institute.csl"
link-citations: true
linestretch: 1.15       
mainfont: Times New Roman
fontsize: 12pt          
geometry: "left=2.54cm,right=2.54cm,top=2.54cm,bottom=2.54cm"
papersize: letter
colorlinks: true
fig-height: 4
fig-width: 7.5
header-includes:
  - \usepackage{threeparttable}
---

{{< include 01-introduction.qmd >}}

{{< include 02-antecedentes.qmd >}}

{{< include 03-methods.qmd >}}

{{< include 04-analisis.qmd >}}

{{< include 05-discussion.qmd >}}

{{< include 06-conclusion.qmd >}}