### 2024-07-19 paper.qmd
Versión final paper.pdf

---
title: "The socialization of meritocracy and market justice preferences at school"
format:
  hikmah-pdf: 
    indent: false
  hikmah-manuscript-pdf: 
    indent: false
author:
  - name: Juan Carlos Castillo
    email: juancastillov@uchile.cl
    orcid: 0000-0003-1265-7854
    affiliations: 
        name: Department of Sociology, Universidad de Chile
        country: Chile
    attributes:
      corresponding: true
  - name: Mauricio Salgado
    email: msalgado@unab.cl
    orcid: 0000-0002-8108-6154
    affiliations: 
        name: Centro de Estudios Públicos, Chile
        country: Chile
  - name: Kevin Carrasco
    email: kevin.carrasco@ug.uchile.cl
    orcid: 0000-0002-9348-8361
    affiliations: 
        name: Center for Social Conflict and Cohesion Studies
        country: Chile
  - name: Andreas Laffert
    email: alaffertt@estudiante.uc.cl
    orcid: 0000-0002-9008-2454
    affiliations: 
        name: Institute of Sociology, Universidad Católica de Chile
        country: Chile
abstract: |
  Previous research has shown that schools justify student performance differences based on meritocratic ideals. However, the link between school meritocracy and students’ justification of inequalities has received little attention. This paper examines how school and family socialization influence students’ beliefs about meritocracy and market-based justifications for access to health, pensions, and education in Chile. Using data from the First Study of Civic Education in Chile, we employ a multilevel modeling strategy to analyze 5,047 students and parents nested within 231 schools. Our findings reveal that a significant proportion of Chilean students agree with market justice principles, at a higher rate than adults. Most students endorse meritocratic views, particularly the reward of effort. Multilevel estimations show strong associations between meritocratic beliefs and market justice preferences: students who believe in the reward of effort and talent are more likely to justify inequalities based on financial capacity. At the school level, market justice preferences are higher in high-status schools but lower in schools with better achievement. Additionally, the conditional influence of meritocracy beliefs weakens in schools with higher socioeconomic status and performance. These results suggest that meritocratic beliefs and market justice preferences begin early and are influenced by the school environment.
keywords: 
  - market justice 
  - meritocracy
  - socialization
  - family
  - schools
  - Chile
bibliography: "input/bib/just-ineq-merit.bib"
csl: "input/bib/multidisciplinary-digital-publishing-institute.csl"
link-citations: true
linestretch: 1.15       
mainfont: Times New Roman
fontsize: 12pt          
geometry: "left=2.54cm,right=2.54cm,top=2.54cm,bottom=2.54cm"
papersize: letter
colorlinks: true
fig-height: 4
fig-width: 7.5
header-includes:
  - \usepackage{threeparttable}
---

This document was last modified at {{< modified >}}
and it was last rendered at {{< now >}}.

{{< include 01-introduction.qmd >}}

{{< include 02-antecedentes.qmd >}}

{{< include 03-methods.qmd >}}

{{< include 04-analisis.qmd >}}

{{< include 05-discussion.qmd >}}

{{< include 06-conclusion.qmd >}}