### 2024-07-22 back_matter.qmd

---
title: "The Socialization of Meritocracy and Market Justice Preferences at School"
subtitle: "Back Matter"
format:
  hikmah-pdf: 
    indent: false
  hikmah-manuscript-pdf: 
    indent: false
bibliography: "input/bib/just-ineq-merit.bib"
csl: "input/bib/multidisciplinary-digital-publishing-institute.csl"
link-citations: true
linestretch: 1.15       
mainfont: Times New Roman
fontsize: 12pt          
geometry: "left=2.54cm,right=2.54cm,top=2.54cm,bottom=2.54cm"
papersize: letter
colorlinks: true
fig-height: 4
fig-width: 7.5
header-includes:
  - \usepackage{threeparttable}
---

# Acknowledgments

The authors acknowledge the support of ANID/FONDECYT grant 1210847 and the Center for Conflict and Social Cohesion Studies (COES) grant ANID/FONDAP 1513009.

# Authors' Contributions

Conceptualization, Juan Carlos Castillo & Mauricio Salgado; Methodology, Juan Carlos Castillo, Kevin Carrasco & Andreas Laffert; Software, Kevin Carrasco & Andreas Laffert; Formal Analysis, Kevin Carrasco & Andreas Laffert; Writing – Original Draft Preparation, Juan Carlos Castillo, Mauricio Salgado, Kevin Carrasco & Andreas Laffert; Visualization, Kevin Carrasco & Andreas Laffert; Supervision, Juan Carlos Castillo.

# Conflict of Interest

The authors report no competing interests to declare.

# Funding

This research received no external funding.

# Data Avaibility Statement

The original data presented in the study are openly available in <https://www.agenciaeducacion.cl/orientar/estudios/estudios-nacionales/>

Access the data, code, and analysis presented in the study are openly available in <https://github.com/educacion-meritocracia/market-justice-preferences-school>.

{{< pagebreak >}}

{{< include 07-supplementary-material.qmd >}}