### 2024-07-22 cover_letter.qmd
Ajustes nuevos
Ajustes finales de:
- Referencias
- Cover Letter
- Funding
- Paper blinded

---
title: "The Socialization of Meritocracy and Market Justice Preferences at School"
subtitle: "Cover Letter"
format:
  hikmah-pdf: 
    indent: false
  hikmah-manuscript-pdf: 
    indent: false
bibliography: "input/bib/just-ineq-merit.bib"
csl: "input/bib/multidisciplinary-digital-publishing-institute.csl"
link-citations: true
linestretch: 1.15       
mainfont: Times New Roman
fontsize: 12pt          
geometry: "left=2.54cm,right=2.54cm,top=2.54cm,bottom=2.54cm"
papersize: letter
colorlinks: true
fig-height: 4
fig-width: 7.5
header-includes:
  - \usepackage{threeparttable}
---

Previous research has shown that schools justify student performance differences based on meritocratic ideals. However, the link between school meritocracy and students’ justification of inequalities has received little attention. This paper examines how school and family socialization influence students’ beliefs about meritocracy and market-based justifications for access to health, pensions, and education in Chile. Using data from the First Study of Civic Education in Chile, we employ a multilevel modeling strategy to analyze 5,047 students and parents nested within 231 schools. Our findings reveal that a significant proportion of Chilean students agree with market justice principles, at a higher rate than adults. Most students endorse meritocratic views, particularly the reward of effort. Multilevel estimations show strong associations between meritocratic beliefs and market justice preferences: students who believe in the reward of effort and talent are more likely to justify inequalities based on financial capacity. At the school level, market justice preferences are higher in high-status schools but lower in schools with better achievement. Additionally, the conditional influence of meritocracy beliefs weakens in schools with higher socioeconomic status and performance. These results suggest that meritocratic beliefs and market justice preferences begin early and are influenced by the school environment.

## Authors Identification

Juan Carlos Castillo, Department of Sociology, Universidad de Chile, Santiago de Chile[^*].  
ORCID: 0000-0003-1265-7854

Mauricio Salgado, School of Sociology, Universidad Andrés Bello.  
ORCID: 0000-0002-8108-6154

Kevin Carrasco, Center for Social Conflict and Cohesion Studies.  
ORCID: 0000-0002-9348-8361

Andreas Laffert, Institute of Sociology, Pontificia Universidad Católica de Chile.  
ORCID: 0000-0002-9008-2454

## Exclusive Submission Statement

The authors confirm that this article has been submitted exclusively to this journal.

[^*]: Corresponding author (juancastillov@uchile.cl - [jc-castillo.com](https://juancarloscastillo.github.io/jc-castillo/) - Ignacio Carrera Pinto 1045 Ñuñoa, Santiago de Chile).