### 2024-07-24 back_matter.qmd

---
title: "The Socialization of Meritocracy and Market Justice Preferences at School"
subtitle: "Supplementary Material"
format:
  hikmah-pdf: 
    indent: false
    keep-tex: true
  hikmah-manuscript-pdf: 
    indent: false
    keep-tex: true
bibliography: "input/bib/just-ineq-merit.bib"
csl: "input/bib/multidisciplinary-digital-publishing-institute.csl"
link-citations: true
linestretch: 1.15       
mainfont: Times New Roman
fontsize: 12pt          
geometry: "left=2.54cm,right=2.54cm,top=2.54cm,bottom=2.54cm"
papersize: letter
colorlinks: true
fig-height: 4
fig-width: 7.5
header-includes:
  - \usepackage{threeparttable}
---

{{< include 07-supplementary-material.qmd >}}