### 2024-08-05 paper.qmd

---
title: "The socialization of meritocracy and market justice preferences at school"
format:
  mdpi-pdf:
    keep-tex: true
mdpi_journal: behavsci
article_type: article
article_status: submit
author:
  - name: Juan Carlos Castillo
    corresponding: true
    orcid: 0000-0003-1265-7854
    email: juancastillov@uchile.cl
    role:
      - Conceptualization
      - Supervision
      - Methodology
      - Writing - review and editing
    affiliations:
      - name: "Universidad de Chile"
        department: Department of Sociology
        id: 1
        country: Chile
  - name: Mauricio Salgado
    orcid: 0000-0002-8108-6154
    role:
      - Conceptualization
      - Methodology
      - Writing - review and editing    
    affiliations:
      - name: Centro de Estudios Públicos
        id: 2
        department: CEP
        country: Chile
  - name: Kevin Carrasco
    orcid: 0000-0002-9348-8361
    role:
      - Methodology
      - formal analysis
      - software
      - visualization
      - Writing - review and editing
    affiliations:
      - name: Center for Social Conflict and Cohesion Studies
        id: 3
        department: COES
        country: Chile
  - name: Andreas Laffert
    orcid: 0000-0002-9008-2454
    role:
      - Methodology
      - formal analysis
      - software
      - visualization
      - Writing - review and editing  
    affiliations:
      - name: Universidad Católica de Chile
        id: 4
        department: Institute of Sociology
        country: Chile
fundingstatement: | 
  The authors acknowledge the support of ANID/FONDECYT grant 1210847 and the Center for Conflict and Social Cohesion Studies (COES) grant ANID/FONDAP 1513009.
dataavailability: |
  The original data presented in the study are openly available in <https://www.agenciaeducacion.cl/orientar/estudios/estudios-nacionales/>
conflictsofinterest: |
  The authors declare no conflict of interest.
abstract: |
  Previous research has shown that schools justify student performance differences based on meritocratic ideals. However, the link between school meritocracy and students’ justification of inequalities has received little attention. This paper examines how school and family socialization influence students’ beliefs about meritocracy and market-based justifications for access to health, pensions, and education in Chile. Using data from the First Study of Civic Education in Chile, we employ a multilevel modeling strategy to analyze 5,047 students and parents nested within 231 schools. Our findings reveal that a significant proportion of Chilean students agree with market justice principles, at a higher rate than adults. Most students endorse meritocratic views, particularly the reward of effort. Multilevel estimations show strong associations between meritocratic beliefs and market justice preferences: students who believe in the reward of effort and talent are more likely to justify inequalities based on financial capacity. At the school level, market justice preferences are higher in high-status schools but lower in schools with better achievement. Additionally, the conditional influence of meritocracy beliefs weakens in schools with higher socioeconomic status and performance. These results suggest that meritocratic beliefs and market justice preferences begin early and are influenced by the school environment.
keywords: [market justice, Meritocracy, socialization, family, schools, Chile]
bibliography: "input/bib/just-ineq-merit.bib"
---

{{< include 01-introduction.qmd >}}

{{< include 02-antecedentes.qmd >}}

{{< include 03-methods.qmd >}}

{{< include 04-analisis.qmd >}}

{{< include 05-discussion.qmd >}}

{{< include 06-conclusion.qmd >}}