### 2024-10-02 paper-blinded.qmd

---
title: "The socialization of meritocracy and market justice preferences at school"
format:
  hikmah-pdf: 
    indent: false
    keep-tex: true
  hikmah-manuscript-pdf: 
    indent: false
    keep-tex: true
abstract: |
  Previous research has shown that schools often justify student performance differences using meritocratic ideals. One potential consequence of such ideals is the legitimization of outcome inequalities across various spheres, including those traditionally associated with equality and redistribution. In this study, we argue that the promotion of meritocratic values during school-age can shape students' beliefs about meritocracy and influence their views on market-based access to health, pensions, and education. Using data from the 2017 National Study of Civic Education in Chile, which includes 5,047 8th-grade students from 231 schools, we estimated a series of multilevel models (lme4 library, R version 4.1.3) to test our hypotheses. Our findings show that a significant proportion of Chilean students agree with market justice principles—more so than adults. Most students endorse meritocratic views, particularly the notion that effort should be rewarded, which strongly correlates with market justice preferences: students who believe in meritocracy are more likely to justify inequalities based on financial capacity. At the school level, market justice preferences are higher in high-status schools but lower in schools with higher academic achievement. Furthermore, the conditional influence of meritocratic beliefs diminishes in schools with higher socioeconomic status and performance levels. These results suggest that the association between meritocratic beliefs and market justice preferences is already established at school age and is shaped by the school environment.
keywords: 
  - market justice 
  - meritocracy
  - socialization
  - family
  - schools
  - Chile
bibliography: "input/bib/just-ineq-merit.bib"
csl: "input/bib/multidisciplinary-digital-publishing-institute.csl"
link-citations: true
linestretch: 1.15       
mainfont: Times New Roman
fontsize: 12pt          
geometry: "left=2.54cm,right=2.54cm,top=2.54cm,bottom=2.54cm"
papersize: letter
colorlinks: true
fig-height: 4
fig-width: 7.5
header-includes:
  - \usepackage{threeparttable}
---
{{< pagebreak >}}

{{< include 01-introduction.qmd >}}

{{< include 02-antecedentes.qmd >}}

{{< include 03-methods.qmd >}}

{{< include 04-analisis.qmd >}}

{{< include 05-discussion.qmd >}}

{{< include 06-conclusion.qmd >}}

{{< pagebreak >}}

{{< include 07-supplementary-material.qmd >}}